---
title: "Run 5 - PH1, PH2 & PH3"
author: "Niamh Errington"
date: "`r format(Sys.time(), '%d %B %Y')`" 
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
    toc_collapsed: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


# Set up

MicroRNAs initially filtered to within the subset used by MiRXES. Chris' LASSO model is then used to transform these, and clustering done on these numbers.

Optimal number of miRNAs: 50

Number of clusters (votes):

* 2 clusters = 4 votes
* 4 clusters = 2 votes
* 5 clusters = 4 votes
* 6 clusters = **5 votes**

Skip to the [Summary](#Summary) section for an overview of the pvalues for each clinical parameter for cluster sizes 2,3 & 6.

3 IDs no Interim / Discovery clarification - set as Interim:
* TB16.0071
* TB12.0464
* TB11.1158

2 miRNA missing from validation set:

```{r, message = FALSE, warning = FALSE}
library(readr)
library(readxl)
library(tidyverse)
library(dunn.test)
library(kableExtra)
library(chisq.posthoc.test)
library(questionr)
library(reshape2)
library(ggpattern)
library(viridis)

#load("Preprocessing.RData")
Clusterphenoall <- read_csv("Clusterpheno.csv") %>% mutate(REVEALgroup = case_when(REVEALscore <7 ~ "Low", REVEALscore %in% c(7,8) ~ "Intermediate", REVEALscore > 8 ~ "High") )
#change smwd
Clusterphenoall <- Clusterphenoall %>% mutate(iswt = ifelse(site == "SHEFFIELD", smwd, NA), smwd = ifelse(site == "SHEFFIELD", NA, smwd))

colnames(Clusterphenoall) <- gsub(x = colnames(Clusterphenoall), "\\.", "-")
colnames(Clusterphenoall)[4:6] <- c("dana1.1", "dana1.1.1", "dana1.1.1.1")

Clusterpheno<- Clusterphenoall %>% filter(PARTITION != "VALIDATION")

clusterfills<- c("deeppink", "dark red", "darksalmon", "yellow", "midnight blue", "cornflowerblue")
```


```{r, eval = FALSE}
sorted_var_microRNAs <- read_csv("sorted_stable_microRNAs_Niamh_new_615s_325m.csv")
sorted_var_microRNAs <- sorted_var_microRNAs %>% mutate(miRnames = gsub("-", ".", X1)) %>% select(-X1, -variance)

# correct pheno from Mirxes data set, D/I taken from t12
pheno <- read_csv("../Phenocleaning/phenotoclean.csv")  
load("../Phenocleaning/miRData.RData") 
new<- new %>% select(SampleID, all_of(sorted_var_microRNAs$miRnames))
miRdata <- new[,colnames(new) %in% c("SampleID", sorted_var_microRNAs$miRnames)]

Run5data <- left_join(pheno, new, by = "SampleID") %>% filter(dana1 %in% c("PH1", "PH2", "PH3"))

write_csv(Run5data, "Run5data.csv")

#write_csv(mirxesMirsfull, "Training.csv")
#write_csv(mirxesval, "Validation.csv")

cluster6 <-read_csv("microRNA_6_clusters_Niamh_new_615s_325m.csv") %>% mutate(cluster6=recode(clusters, `1`="A", `2`="B",`3`="C", `4`="D", `5`="E", `6`="F"))  %>% select(-clusters)

cluster2<- read_csv("microRNA_2_clusters_Niamh_new_615s_325m.csv") %>% mutate(cluster2=recode(clusters, `1`="A", `2`="B"))

Clusterpheno <- left_join(pheno, miRdata, by = "SampleID") %>% filter(dana1 %in% c("PH1", "PH2", "PH3")) %>% left_join(., cluster6, by = "SampleID")

write_csv(Clusterpheno, "Clusterpheno.csv")
```

# Clinical Associations

Each parameter is assessed for normality, then an appropriate statistical test applied to look for associations to cluster membership. Where the variances are unequal, the data are log transformed. Where an inital p value is significant, post hoc tests are applied to look between groups. 

```{r}
BMI <- c(Median = median(Clusterpheno$bmi, na.rm = TRUE), IQR = IQR(Clusterpheno$bmi, na.rm = TRUE))
age <- c(Median = median(Clusterpheno$age, na.rm = TRUE), IQR = IQR(Clusterpheno$age, na.rm = TRUE))
NTproBNP <- c(Median = median(Clusterpheno$ntprobnp, na.rm = TRUE), IQR = IQR(Clusterpheno$ntprobnp, na.rm = TRUE))
`Uric Acid` <- c(Median = median(Clusterpheno$ua, na.rm = TRUE), IQR = IQR(Clusterpheno$ua, na.rm = TRUE))
`6MWD` <- c(Median = median(Clusterpheno$smwd, na.rm = TRUE), IQR = IQR(Clusterpheno$smwd, na.rm = TRUE))
ISWT <- c(Median = median(Clusterpheno$iswt, na.rm = TRUE), IQR = IQR(Clusterpheno$iswt, na.rm = TRUE))

mPAP <- c(Median = median(Clusterpheno$mPAP, na.rm = TRUE), IQR = IQR(Clusterpheno$mPAP, na.rm = TRUE))
sPAP <- c(Median = median(Clusterpheno$sPAP, na.rm = TRUE), IQR = IQR(Clusterpheno$sPAP, na.rm = TRUE))
dPAP <- c(Median = median(Clusterpheno$dPAP, na.rm = TRUE), IQR = IQR(Clusterpheno$dPAP, na.rm = TRUE))
mRAP <- c(Median = median(Clusterpheno$mRAP, na.rm = TRUE), IQR = IQR(Clusterpheno$mRAP, na.rm = TRUE))
SvO2 <- c(Median = median(Clusterpheno$SvO2, na.rm = TRUE), IQR = IQR(Clusterpheno$SvO2, na.rm = TRUE))
PVR <- c(Median = median(Clusterpheno$PVRdynes, na.rm = TRUE), IQR = IQR(Clusterpheno$PVRdynes, na.rm = TRUE))
PAWP <- c(Median = median(Clusterpheno$PAWP, na.rm = TRUE), IQR = IQR(Clusterpheno$PAWP, na.rm = TRUE))
BPDIA <- c(Median = median(Clusterpheno$bpdia, na.rm = TRUE), IQR = IQR(Clusterpheno$bpdia, na.rm = TRUE))
BPSYS <- c(Median = median(Clusterpheno$bpsys, na.rm = TRUE), IQR = IQR(Clusterpheno$bpsys, na.rm = TRUE))
eGFR <- c(Median = median(Clusterpheno$eGFR, na.rm = TRUE), IQR = IQR(Clusterpheno$eGFR, na.rm = TRUE))
`Cardiac Output` <- c(Median = median(Clusterpheno$CardiacOutput, na.rm = TRUE), IQR = IQR(Clusterpheno$CardiacOutput, na.rm = TRUE))
`Cardiac Index` <- c(Median = median(Clusterpheno$CardiacIndex, na.rm = TRUE), IQR = IQR(Clusterpheno$CardiacIndex, na.rm = TRUE))
Creatinine <- c(Median = median(Clusterpheno$Creatinine, na.rm = TRUE), IQR = IQR(Clusterpheno$Creatinine, na.rm = TRUE))
`Platelet Count` <- c(Median = median(Clusterpheno$platelet_count, na.rm = TRUE), IQR = IQR(Clusterpheno$platelet_count, na.rm = TRUE))
`TLCO predicted` <- c(Median = median(Clusterpheno$tlco_predicted, na.rm = TRUE), IQR = IQR(Clusterpheno$tlco_predicted, na.rm = TRUE))

FVCP <- c(Mean = mean(Clusterpheno$fvcp, na.rm = TRUE), SD = sd(Clusterpheno$fvcp, na.rm = TRUE))
kable(rbind(BMI, age, NTproBNP, `Uric Acid`, `6MWD`, ISWT, PVR, mPAP, sPAP, dPAP, mRAP, SvO2, BPDIA, BPSYS, PAWP, eGFR, `Cardiac Output`, `Cardiac Index`,`Platelet Count`, Creatinine, `TLCO predicted`), caption = "Not normally distributed variables") %>% kable_styling(full_width = TRUE)

kable(FVCP, col.names = "FVCP", caption = "Normally distributed variables") %>% kable_styling(full_width = TRUE)
```



## PH Status

```{r}
kable(addmargins(table(Clusterpheno$dana, Clusterpheno$cluster6)), caption = "PH status breakdown per cluster") %>% kable_styling(full_width = TRUE)

kable(prop.table(table(Clusterpheno$dana1, Clusterpheno$cluster6),2)*100, caption = "% composition dana 1 per cluster") %>% kable_styling(full_width = TRUE)

kable(prop.table(table(Clusterpheno$dana, Clusterpheno$cluster6),2)*100, caption = "% composition per cluster") %>% kable_styling(full_width = TRUE)
fisher.test(table(Clusterpheno$dana, Clusterpheno$cluster6), conf.level = 0.95, simulate.p.value=TRUE)

#Posthoc Chisq
kable(chisq.posthoc.test(table(Clusterpheno$dana, Clusterpheno$cluster6), method = "BH"))  %>% kable_styling(full_width = TRUE)
#as.data.frame(c2DANA) %>% filter(Var2 == "A") %>% select(-Var2) %>% ggplot(aes(x="", y=Freq, fill=Var1)) + geom_bar(stat="identity", width = 1, color = "white") + coord_polar("y", start =0) + theme_void() + scale_fill_brewer(palette = "Set1") + ggtitle("A")
cp <- coord_polar(theta = "y")
cp$is_free <- function() TRUE

Clusterpheno$danalong <- recode(Clusterpheno$dana, "PH1.1" = "Idiopathic PAH", "PH1.2" = "Heritable PAH", "PH1.3" = "Drug and toxin induced PAH", "PH1.4" = "Associated PAH", "PH2" = "PH-LHD", "PH3" = "PH-lung")
Clusterpheno <- Clusterpheno%>% mutate(danalong = factor(danalong, levels = c("Idiopathic PAH", "Heritable PAH", "Drug and toxin induced PAH","Associated PAH","PH-LHD","PH-lung")))

as.data.frame(table(Clusterpheno$danalong, Clusterpheno$cluster6)) %>% group_by(Var2, Var1) %>% ggplot(aes(x="", y=Freq, fill=Var1)) + geom_bar(stat="identity", width = 1, color = "white") + cp + theme_void()  + facet_wrap(~Var2, scales = "free") + theme(aspect.ratio = 1) + scale_fill_manual(values = clusterfills) +  guides(fill=guide_legend(title=NULL))

as.data.frame(table(Clusterpheno$danalong, Clusterpheno$cluster6))  %>% group_by(Var2, Var1) %>% ggplot(aes(x=Var2, y=Freq, fill=Var1)) + geom_bar(stat="identity")  + scale_fill_manual(values = clusterfills) + ggtitle("PH Status, 6 Clusters") + labs(x = "Cluster", y = "", size = 14) +  guides(fill=guide_legend(title=NULL)) + theme(panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey"), strip.text = element_text(size = 14)) +  guides(fill=guide_legend(title=NULL))

Clusterpheno$dana1long <- recode(Clusterpheno$dana1, "PH1" = "PAH", "PH2" = "PH-LHD", "PH3" = "PH-lung")

as.data.frame(table(Clusterpheno$dana1long, Clusterpheno$cluster6))  %>% group_by(Var2, Var1) %>% ggplot(aes(x=Var2, y=Freq, fill=Var1)) + geom_bar(stat="identity")  + scale_fill_manual(values = c("deeppink", "midnight blue", "cornflowerblue")) + ggtitle("PH Status, 6 Clusters") + labs(x = "Cluster", y = "") +  guides(fill=guide_legend(title=NULL)) + theme(panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey")) +  guides(fill=guide_legend(title=NULL))

as.data.frame(table(Clusterpheno$dana1long, Clusterpheno$cluster6)) %>% group_by(Var2, Var1) %>% ggplot(aes(x="", y=Freq, fill=Var1)) + geom_bar(stat="identity", width = 1, color = "white") + cp + theme_void()  + facet_wrap(~Var2, scales = "free") + theme(aspect.ratio = 1, text = element_text(size = 14)) + scale_fill_manual(values = c("deeppink", "midnight blue", "cornflowerblue")) +  guides(fill=guide_legend(title=NULL))
```

## Sex

```{r}
kable(addmargins(table(Clusterpheno$sex, Clusterpheno$cluster6))) %>% kable_styling(full_width = TRUE)
sex6<- chisq.test(table(Clusterpheno$sex, Clusterpheno$cluster6))
sex6

library(ggpattern)
as.data.frame(table(Clusterpheno$cluster6, Clusterpheno$sex)) %>% ggplot(aes(x = Var1, y = Freq, fill = Var1, pattern = Var2)) + scale_fill_manual(values = clusterfills)  + geom_bar_pattern(stat = "identity", color = "black", 
  pattern_fill = "black",
                   pattern_angle = 45,
                   pattern_density = 0.1,
                   pattern_spacing = 0.025,) + scale_pattern_manual(values = c(M = "stripe", F = "none")) +
  ggtitle("Sex") + labs(x = "Cluster", y = "") + guides(fill=guide_legend(title=NULL)) + theme(panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey"))


as.data.frame(table(Clusterpheno$cluster6, Clusterpheno$sex)) %>% ggplot(aes(x = Var1, y = Freq, fill = Var2)) + scale_fill_manual(values = c("midnight blue", "cornflowerblue"))  + geom_bar(stat = "identity") + ggtitle("Sex") + labs(x = "Cluster", y = "") + guides(fill=guide_legend(title=NULL)) + theme(panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey"))


as.data.frame(table(Clusterpheno$dana, Clusterpheno$sex)) %>% ggplot(aes(x = Var1, y = Freq, fill = Var2)) + geom_bar(stat = "identity") + scale_fill_manual(values = c("midnight blue", "cornflowerblue")) + ggtitle("PH status by sex") + labs(x = "Cluster", y = "") + guides(fill=guide_legend(title=NULL)) + theme(panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey"))
```

## FC
```{r, warning = FALSE, message = FALSE}
kable(cbind(table(is.na(Clusterpheno$fclass)), prop.table(table(is.na(Clusterpheno$fclass)))*100), col.names = c("Missing Values", "%")) %>% kable_styling(full_width = TRUE)
```

```{r}
kable(addmargins(table(Clusterpheno$fclass,  Clusterpheno$cluster6))) %>% kable_styling(full_width = TRUE)
kable(prop.table(table(Clusterpheno$fclass, Clusterpheno$cluster6),2)*100, caption = "% composition per cluster") %>% kable_styling(full_width = TRUE)
fc6<- fisher.test(table(Clusterpheno$fclass, Clusterpheno$cluster6), conf.level = 0.95, simulate.p.value=TRUE)
fc6

as.data.frame(table(Clusterpheno$cluster6, Clusterpheno$fclass)) %>% ggplot(aes(x = Var1, y = Freq, fill = Var1, pattern = Var2)) + scale_fill_manual(values = clusterfills)  + geom_bar_pattern(stat = "identity", color = "black", 
  pattern_fill = "black",
                   pattern_angle = 45,
                   pattern_density = 0.1,
                   pattern_spacing = 0.025,) + scale_pattern_manual(values = c(`1` = "stripe", `2` = "none", `3` = "circle", `4` = "crosshatch")) +
  ggtitle("Functional class") + labs(x = "Cluster", y = "") + guides(fill=guide_legend(title=NULL)) + theme(panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey"))


as.data.frame(table(Clusterpheno$cluster6, Clusterpheno$fclass)) %>% ggplot(aes(x = Var1, y = Freq, fill = Var2)) + geom_bar(stat = "identity") + scale_fill_manual(values = c("deeppink","yellow", "midnight blue", "cornflowerblue")) + ggtitle("Functional Class") + labs(x = "Cluster", y = "n") + guides(fill=guide_legend(title=NULL)) + theme(panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey"), text = element_text(size = 14))
```


## REVEAL group


```{r, warning = FALSE, message = FALSE}
kable(cbind(table(is.na(Clusterpheno$REVEALgroup)), prop.table(table(is.na(Clusterpheno)))*100), col.names = c("Missing Values", "%")) %>% kable_styling(full_width = TRUE)
```

```{r}
kable(addmargins(table(Clusterpheno$REVEALgroup,  Clusterpheno$cluster6))) %>% kable_styling(full_width = TRUE)
kable(prop.table(table(Clusterpheno$REVEALgroup, Clusterpheno$cluster6),2)*100, caption = "% composition per cluster") %>% kable_styling(full_width = TRUE)
Reveal6<- fisher.test(table(Clusterpheno$REVEALgroup, Clusterpheno$cluster6), conf.level = 0.95, simulate.p.value=TRUE)
Reveal6

kable(chisq.posthoc.test(table(Clusterpheno$REVEALgroup, Clusterpheno$cluster6), method = "BH"))  %>% kable_styling(full_width = TRUE)

as.data.frame(table(Clusterpheno$cluster6, Clusterpheno$REVEALgroup)) %>% ggplot(aes(x = Var1, y = Freq, fill = Var1, pattern = Var2)) + scale_fill_manual(values = c("deeppink", "dark red", "darksalmon", "yellow", "midnight blue", "cornflowerblue"))  + geom_bar_pattern(stat = "identity", color = "black", 
  pattern_fill = "black",
                   pattern_angle = 45,
                   pattern_density = 0.1,
                   pattern_spacing = 0.025,) + scale_pattern_manual(values = c(Low = "stripe", Intermediate = "circle", High = "none")) +
  ggtitle("REVEAL score risk group") + labs(x = "Cluster", y = "") + guides(fill=guide_legend(title=NULL)) + theme(panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey"))

as.data.frame(table(Clusterpheno$cluster6, Clusterpheno$REVEALgroup)) %>% ggplot(aes(x = Var1, y = Freq, fill = Var2)) + geom_bar(stat = "identity") + scale_fill_manual(values = c("deeppink", "midnight blue", "cornflowerblue")) + ggtitle("REVEAL score risk group") + labs(x = "Cluster", y = "") + guides(fill=guide_legend(title=NULL)) + theme(panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey"), text = element_text(size = 14))
```

## Site

Fisher test (simulate p values) p val < 0.05

```{r}
kable(addmargins(table(Clusterpheno$site,  Clusterpheno$cluster6))) %>% kable_styling(full_width = TRUE)
site6<- fisher.test(table(Clusterpheno$site, Clusterpheno$cluster6), conf.level = 0.95, simulate.p.value=TRUE)
site6
#Posthoc Chisq
kable(chisq.posthoc.test(table(Clusterpheno$site, Clusterpheno$cluster6), method = "BH"))  %>% kable_styling(full_width = TRUE)

as.data.frame(table(Clusterpheno$cluster6, Clusterpheno$site)) %>% ggplot(aes(x = Var1, y = Freq, fill = Var1, pattern = Var2)) + scale_fill_manual(values = clusterfills)  + geom_bar_pattern(stat = "identity", color = "black", 
  pattern_fill = "black",
                   pattern_angle = 45,
                   pattern_density = 0.1,
                   pattern_spacing = 0.025,) + scale_pattern_manual(values = c(SHEFFIELD = "stripe", IMPERIAL = "circle", CAMBRIDGE = "none")) +
  ggtitle("Site") + labs(x = "Cluster", y = "") + guides(fill=guide_legend(title=NULL)) + theme(panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey"))

as.data.frame(table(Clusterpheno$cluster6, Clusterpheno$site)) %>% ggplot(aes(x = Var1, y = Freq, fill = Var2)) + geom_bar(stat = "identity") + scale_fill_manual(values = c("deeppink","cornflowerblue", "midnight blue")) + ggtitle("Site") + labs(x = "Cluster", y = "") + guides(fill=guide_legend(title=NULL))
```

## BMI

Not normally distributed. Large variance ==> log transformed

```{r, warning = FALSE, message = FALSE}
kable(cbind(table(is.na(Clusterpheno$bmi)), prop.table(table(is.na(Clusterpheno$bmi)))*100), col.names = c("Missing Values", "%")) %>% kable_styling(full_width = TRUE)
```

```{r, warning = FALSE, message = FALSE}
Clusterpheno %>% select(cluster6, bmi) %>% filter(!is.na(bmi)) %>% ggplot(aes(x= cluster6, y =bmi)) + geom_boxplot(fill=clusterfills) + labs(x = "Cluster", y = "BMI") + theme(panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey"))


Clusterpheno %>% filter(!is.na(bmi)) %>% group_by(cluster6) %>% summarise(Shapiro.pvalue = shapiro.test(bmi)$p.value, Mean = mean(bmi), Variance = var(bmi), LogMean = mean(log(bmi)), LogVariance = var(log(bmi)), Median = median(bmi), IQR = IQR(bmi)) %>% kable(.) %>% kable_styling(full_width = TRUE)

bmi6<- kruskal.test(log(bmi) ~ cluster6, data = Clusterpheno)
bmi6
```

## Age

Not normally distributed. Large variances ==> log transformed

```{r, warning = FALSE, message = FALSE}
kable(cbind(table(is.na(Clusterpheno$age)), prop.table(table(is.na(Clusterpheno$age)))*100), col.names = c("Missing Values", "%")) %>% kable_styling(full_width = TRUE)

Clusterpheno %>% select(cluster6, age) %>% filter(!is.na(age)) %>% ggplot(aes(x= cluster6, y =age)) + geom_boxplot(fill=clusterfills) + labs(x = "Cluster", y = "Age") + theme(panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey"))

Clusterpheno %>%filter(!is.na(age)) %>% group_by(cluster6) %>% summarise(Shapiro.pvalue = shapiro.test(age)$p.value, Mean =mean(age), Variance = var(age), LogMean = mean(log(age)), LogVariance = var(log(age)), Median = median(age), IQR = IQR(age)) %>% kable(.) %>% kable_styling(full_width = TRUE)

age6<- kruskal.test(log(age) ~ cluster6, data = Clusterpheno)
age6
```

## NTproBNP

Not all normally distributed.

```{r, warning = FALSE, message = FALSE}
kable(cbind(table(is.na(Clusterpheno$ntprobnp)), prop.table(table(is.na(Clusterpheno$ntprobnp)))*100), col.names = c("Missing Values", "%")) %>% kable_styling(full_width = TRUE)
Clusterpheno %>% select(cluster6, ntprobnp) %>% ggplot(aes(x= cluster6, y = ntprobnp)) + geom_boxplot(fill=clusterfills) + labs(x = "Cluster", y = "NT-proBNP") + theme(panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey"))

group_by(Clusterpheno, cluster6) %>% summarise(Shapiro.pvalue = shapiro.test(ntprobnp)$p.value, Variance = var(ntprobnp), Mean = mean(ntprobnp), Median = median(ntprobnp), IQR = IQR(ntprobnp)) %>% kable(.) %>% kable_styling(full_width = TRUE)

ntprobnp6<- kruskal.test(ntprobnp ~ cluster6, data = Clusterpheno)
ntprobnp6
dunn.test(Clusterpheno$ntprobnp, Clusterpheno$cluster6, method = "BH")
```

## Uric Acid

Not all normally distributed.

```{r, warning = FALSE, message = FALSE}
kable(table(is.na(Clusterpheno$ua)), col.names = c("Missing Values", ""))

Clusterpheno %>% select(cluster6, ua) %>% ggplot(aes(x= cluster6, y = ua)) + geom_boxplot(fill=clusterfills) + labs(x = "Cluster", y = "Uric Acid") + theme(panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey"))

group_by(Clusterpheno, cluster6) %>% summarise(Shapiro.pvalue = shapiro.test(ua)$p.value, Variance = var(ua), Mean = mean(ua), Median = median(ua), IQR = IQR(ua)) %>% kable(.) %>% kable_styling(full_width = TRUE)

ua6<- kruskal.test(ua ~ cluster6, data = Clusterpheno)
ua6
dunn.test(Clusterpheno$ua, Clusterpheno$cluster6, method = "BH")
```

## FVCP

Normally distributed. Large variance ==> log transformed

```{r, warning = FALSE, message = FALSE}
kable(cbind(table(is.na(Clusterpheno$fvcp)), prop.table(table(is.na(Clusterpheno$fvcp)))*100), col.names = c("Missing Values", "%")) %>% kable_styling(full_width = TRUE)
Clusterpheno %>% select(cluster6, fvcp) %>% ggplot(aes(x= cluster6, y = fvcp)) + geom_boxplot(fill=clusterfills) + labs(x = "Cluster", y = "FVCP") + theme(panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey"))

Clusterpheno %>% filter(!is.na(fvcp)) %>% group_by(cluster6) %>% summarise(Shapiro.pvalue = shapiro.test(fvcp)$p.value, Variance = var(fvcp), Mean = mean(fvcp), SD = sd(fvcp), LogMean = mean(log(fvcp)), LogVariance = var(log(fvcp))) %>% kable(.) %>% kable_styling(full_width = TRUE)

fvcp6<- aov(log(fvcp) ~ cluster6, data = Clusterpheno)
summary(fvcp6)
```

## Treatment 

```{r, warning = FALSE, message = FALSE}
kable(addmargins(table(Clusterpheno$treated, Clusterpheno$cluster6))) %>% kable_styling(full_width = TRUE)
kable(prop.table(table(Clusterpheno$treated, Clusterpheno$cluster6),2)*100) %>% kable_styling(full_width = TRUE)
treatment6<- chisq.test(table(Clusterpheno$treated, Clusterpheno$cluster6))
treatment6
as.data.frame(table(Clusterpheno$cluster6, Clusterpheno$treated)) %>% ggplot(aes(x = Var1, y = Freq, fill = Var2)) + geom_bar(stat = "identity") + scale_fill_manual(values = c("tomato2","dark red")) + ggtitle("Treated") + labs(x = "Cluster", y = "") + guides(fill=guide_legend(title=NULL))

#for posthoc Chisq
kable(chisq.posthoc.test(table(Clusterpheno$treated, Clusterpheno$cluster6), method = "BH"))  %>% kable_styling(full_width = TRUE)
```

```{r}
as.data.frame(table(Clusterpheno$dana1, Clusterpheno$treated)) %>% ggplot(aes(x = Var1, y = Freq, fill = Var2)) + geom_bar(stat = "identity") + scale_fill_manual(values = c("tomato2","dark red")) + ggtitle("Treatment breakdown by PH group") + theme_minimal() + theme(axis.text.x = element_text(angle = 90)) + labs(x = "Cluster", y = "") + guides(fill=guide_legend(title=NULL))
```

```{r}
as.data.frame(table(Clusterpheno$site, Clusterpheno$treated)) %>% ggplot(aes(x = Var1, y = Freq, fill = Var2)) + geom_bar(stat = "identity") + scale_fill_manual(values = c("tomato2","dark red")) + ggtitle("Treatment breakdown by site") + theme_minimal() + theme(axis.text.x = element_text(angle = 90)) + labs(x = "Cluster", y = "") + guides(fill=guide_legend(title=NULL))
```



### Treatment breakdown

```{r}
kable(addmargins(table(Clusterpheno$ERA, Clusterpheno$cluster6)), caption = "ERAs") %>% kable_styling(full_width = TRUE)
kable(addmargins(table(Clusterpheno$Prostanoid, Clusterpheno$cluster6)), caption = "Prostanoids") %>% kable_styling(full_width = TRUE)
kable(addmargins(table(Clusterpheno$Other, Clusterpheno$cluster6)), caption = "Other PH drugs") %>% kable_styling(full_width = TRUE)
```

## 6MWD

Not normally distributed, large variance ==> log transformed

```{r, warning = FALSE, message = FALSE}
kable(cbind(table(is.na(Clusterpheno$smwd)), prop.table(table(is.na(Clusterpheno$smwd)))*100), col.names = c("Missing Values", "%")) %>% kable_styling(full_width = TRUE)
Clusterpheno %>% select(cluster6, smwd) %>% ggplot(aes(x= cluster6, y =smwd)) + geom_boxplot(fill=clusterfills) + labs(x = "Cluster", y = "6MWD") + theme(panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey"))

Clusterpheno %>% filter(!is.na(smwd)) %>% group_by(cluster6) %>% summarise(Shapiro.pvalue = shapiro.test(smwd)$p.value, Variance = var(smwd), Mean = mean(smwd), logVariance = var(log(smwd)), logmean = mean(log(smwd)), Median = median(smwd), IQR = IQR(smwd)) %>% kable(.) %>% kable_styling(full_width = TRUE)

smwd6<- kruskal.test(log(smwd) ~ cluster6, data = Clusterpheno)
smwd6
dunn.test(log(Clusterpheno$smwd), Clusterpheno$cluster6, method = "BH")

```


## ISWT

Sheffield only

Not normally distributed, log transformed

```{r, warning = FALSE, message = FALSE}
kable(cbind(table(is.na(Clusterpheno$iswt)), prop.table(table(is.na(Clusterpheno$iswt)))*100), col.names = c("Missing Values", "%")) %>% kable_styling(full_width = TRUE)
Clusterpheno %>% select(cluster6, iswt) %>% ggplot(aes(x= cluster6, y =iswt)) + geom_boxplot(fill=clusterfills) + labs(x = "Cluster", y = "6MWD") + theme(panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey"))

Clusterpheno %>% filter(!is.na(iswt)) %>% group_by(cluster6) %>% summarise(Shapiro.pvalue = shapiro.test(iswt)$p.value, Variance = var(iswt), Mean = mean(iswt), logVariance = var(log(iswt)), logmean = mean(log(iswt)), Median = median(iswt), IQR = IQR(iswt)) %>% kable(.) %>% kable_styling(full_width = TRUE)

iswt6<- kruskal.test(log(iswt) ~ cluster6, data = Clusterpheno)
iswt6
dunn.test(log(Clusterpheno$iswt), Clusterpheno$cluster6, method = "BH")

```


## BPDIA

Not all normally distributed, large variance ==> log transformed

```{r, warning = FALSE, message = FALSE}
kable(cbind(table(is.na(Clusterpheno$bpdia)), prop.table(table(is.na(Clusterpheno$bpdia)))*100), col.names = c("Missing Values", "%")) %>% kable_styling(full_width = TRUE)
Clusterpheno %>% select(cluster6, bpdia) %>% ggplot(aes(x= cluster6, y =bpdia)) + geom_boxplot(fill=clusterfills) + labs(x = "Cluster", y = "BPDIA") + theme(panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey"))

Clusterpheno %>% filter(!is.na(bpdia)) %>% group_by(cluster6) %>% summarise(Shapiro.pvalue = shapiro.test(bpdia)$p.value, Variance = var(bpdia), Mean = mean(bpdia), logVariance = var(log(bpdia)), logmean = mean(log(bpdia)), Median = median(bpdia), IQR = IQR(bpdia)) %>% kable(.) %>% kable_styling(full_width = TRUE)

bpdia6<- kruskal.test(log(bpdia) ~ cluster6, data = Clusterpheno)
bpdia6
```

## BPSYS

Not normally distributed, large variance ==> log transformed

```{r, warning = FALSE, message = FALSE}
kable(cbind(table(is.na(Clusterpheno$bpsys)), prop.table(table(is.na(Clusterpheno$bpsys)))*100), col.names = c("Missing Values", "%")) %>% kable_styling(full_width = TRUE)
Clusterpheno %>% select(cluster6, bpsys) %>% ggplot(aes(x= cluster6, y =bpsys)) + geom_boxplot(fill=clusterfills) + labs(x = "Cluster", y = "BPSYS") + theme(panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey"))

Clusterpheno %>% filter(!is.na(bpsys)) %>% group_by(cluster6) %>% summarise(Shapiro.pvalue = shapiro.test(bpsys)$p.value, Variance = var(bpsys), Mean = mean(bpsys), logVariance = var(log(bpsys)), logmean = mean(log(bpsys)), Median = median(bpsys), IQR = IQR(bpsys))  %>% kable(.) %>% kable_styling(full_width = TRUE)

bpsys6<- kruskal.test(log(bpsys) ~ cluster6, data = Clusterpheno)
bpsys6
```


## PH sub status

```{r, warning = FALSE, message = FALSE}
kable(addmargins(table(Clusterpheno$dana1.1.1, Clusterpheno$cluster6))) %>% kable_styling(full_width = TRUE)
fisher.test(table(Clusterpheno$dana1.1.1, Clusterpheno$cluster6), conf.level = 0.95, simulate.p.value=TRUE)

as.data.frame(table(Clusterpheno$dana1.1.1, Clusterpheno$cluster6))  %>% group_by(Var2, Var1) %>% ggplot(aes(x=Var2, y=Freq, fill=Var1)) + geom_bar(stat="identity") + ggtitle("PH Status, 6 Clusters") + labs(x = "Cluster", y = "") + guides(fill=guide_legend(title=NULL))

#+ scale_fill_manual(values = c("tomato2","dark red", "red", "darksalmon", "midnight blue", "cornflowerblue")) 

PH1.4 <- Clusterpheno %>% filter(str_detect(dana1.1.1, "^PH1.4"))
as.data.frame(table(PH1.4$dana1.1.1, PH1.4$cluster6))  %>% group_by(Var2, Var1) %>% ggplot(aes(x=Var2, y=Freq, fill=Var1)) + geom_bar(stat="identity") + ggtitle("PH1.4 Status, 6 Clusters") + labs(x = "Cluster", y = "") + guides(fill=guide_legend(title=NULL)) + scale_fill_manual(values = c("deeppink","dark red", "midnight blue", "cornflowerblue")) 
```

## PVR

Not normally distributed, unequal variance ==> log transformed

```{r, warning = FALSE, message = FALSE}
kable(cbind(table(is.na(Clusterpheno$PVRdynes)), prop.table(table(is.na(Clusterpheno$PVRdynes)))*100), col.names = c("Missing Values", "%")) %>% kable_styling(full_width = TRUE)
Clusterpheno %>% select(cluster6, PVRdynes) %>% filter(!is.na(PVRdynes)) %>% group_by(cluster6) %>% summarise(Shapiro.pvalue = shapiro.test(PVRdynes)$p.value, Mean = mean(PVRdynes), logmean = mean(log(PVRdynes)), Variance = var(PVRdynes), logvar = var(log(PVRdynes)), min = min(PVRdynes), max =  max(PVRdynes), median = median(PVRdynes), IQR = IQR(PVRdynes)) %>% kable(.) %>% kable_styling(full_width = TRUE)

Clusterpheno %>% select(cluster6, PVRdynes) %>% filter(!is.na(PVRdynes)) %>% ggplot(aes(x= cluster6, y =PVRdynes)) + geom_boxplot(fill=clusterfills) + labs(x = "Cluster", y = "PVR (Dynes)") + theme(panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey"))

pvr6<- kruskal.test(log(PVRdynes) ~ cluster6, data = Clusterpheno)
pvr6
dunn.test(log(Clusterpheno$PVRdynes), Clusterpheno$cluster6, method = "BH")
```

## mPAP

Not normally distributed, unequal variance ==> log transformed

```{r, warning = FALSE, message = FALSE}
kable(cbind(table(is.na(Clusterpheno$mPAP)), prop.table(table(is.na(Clusterpheno$mPAP)))*100), col.names = c("Missing Values", "%")) %>% kable_styling(full_width = TRUE)
Clusterpheno %>% select(cluster6, mPAP) %>% filter(!is.na(mPAP)) %>% group_by(cluster6) %>% summarise(Shapiro.pvalue = shapiro.test(mPAP)$p.value, Mean = mean(mPAP), logmean = mean(log(mPAP)), Variance = var(mPAP), logvar = var(log(mPAP)), min = min(mPAP), max =  max(mPAP), median = median(mPAP), IQR = IQR(mPAP))%>% kable(.) %>% kable_styling(full_width = TRUE)

Clusterpheno %>% select(cluster6, mPAP) %>% filter(!is.na(mPAP)) %>% ggplot(aes(x= cluster6, y =mPAP)) + geom_boxplot(fill=clusterfills) + labs(x = "Cluster", y = "mPAP (mm Hg)") + theme(panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey"))

mpap6<- kruskal.test(log(mPAP) ~ cluster6, data = Clusterpheno)
mpap6
dunn.test(log(Clusterpheno$mPAP), Clusterpheno$cluster6, method = "BH")
```

## sPAP

Not normally distributed, unequal variance ==> log transformed

```{r, warning = FALSE, message = FALSE}
kable(cbind(table(is.na(Clusterpheno$sPAP)), prop.table(table(is.na(Clusterpheno$sPAP)))*100), col.names = c("Missing Values", "%")) %>% kable_styling(full_width = TRUE)
Clusterpheno %>% select(cluster6, sPAP) %>% filter(!is.na(sPAP)) %>% group_by(cluster6) %>% summarise(Shapiro.pvalue = shapiro.test(sPAP)$p.value, Mean = mean(sPAP), logmean = mean(log(sPAP)), Variance = var(sPAP), logvar = var(log(sPAP)), min = min(sPAP), max =  max(sPAP), median = median(sPAP), IQR = IQR(sPAP))%>% kable(.) %>% kable_styling(full_width = TRUE)

Clusterpheno %>% select(cluster6, sPAP) %>% filter(!is.na(sPAP)) %>% ggplot(aes(x= cluster6, y =sPAP)) + geom_boxplot(fill=clusterfills) + labs(x = "Cluster", y = "sPAP (mm Hg)") + theme(panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey"))

spap6<- kruskal.test(log(sPAP) ~ cluster6, data = Clusterpheno)
spap6
dunn.test(log(Clusterpheno$sPAP), Clusterpheno$cluster6, method = "BH")
```

## dPAP

Not normally distributed, unequal variance ==> log transformed

```{r, warning = FALSE, message = FALSE}
kable(cbind(table(is.na(Clusterpheno$dPAP)), prop.table(table(is.na(Clusterpheno$dPAP)))*100), col.names = c("Missing Values", "%")) %>% kable_styling(full_width = TRUE)
Clusterpheno %>% select(cluster6, dPAP) %>% filter(!is.na(dPAP)) %>% group_by(cluster6) %>% summarise(Shapiro.pvalue = shapiro.test(dPAP)$p.value, Mean = mean(dPAP), logmean = mean(log(dPAP)), Variance = var(dPAP), logvar = var(log(dPAP)), min = min(dPAP), max =  max(dPAP), median = median(dPAP), IQR = IQR(dPAP))%>% kable(.) %>% kable_styling(full_width = TRUE)

Clusterpheno %>% select(cluster6, dPAP) %>% filter(!is.na(dPAP)) %>% ggplot(aes(x= cluster6, y =dPAP)) + geom_boxplot(fill=clusterfills) + labs(x = "Cluster", y = "dPAP (mm Hg)") + theme(panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey"))

dpap6<- kruskal.test(log(dPAP) ~ cluster6, data = Clusterpheno)
dpap6
dunn.test(log(Clusterpheno$dPAP), Clusterpheno$cluster6, method = "BH")
```

## mRAP

Not normally distributed, log transformed

```{r, warning = FALSE, message = FALSE}
kable(cbind(table(is.na(Clusterpheno$mRAP)), prop.table(table(is.na(Clusterpheno$mRAP)))*100), col.names = c("Missing Values", "%")) %>% kable_styling(full_width = TRUE)
Clusterpheno %>% select(cluster6, mRAP) %>% filter(!is.na(mRAP)) %>% group_by(cluster6) %>% summarise(Shapiro.pvalue = shapiro.test(mRAP)$p.value, Mean = mean(mRAP), logmean = mean(log(mRAP)), Variance = var(mRAP), logvar = var(log(mRAP)), min = min(mRAP), max =  max(mRAP), median = median(mRAP), IQR = IQR(mRAP))%>% kable(.) %>% kable_styling(full_width = TRUE)

Clusterpheno %>% select(cluster6, mRAP) %>% filter(!is.na(mRAP)) %>% ggplot(aes(x= cluster6, y =mRAP)) + geom_boxplot(fill=clusterfills) + labs(x = "Cluster", y = "mRAP (mm Hg)") + theme(panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey"))

mRAP6<- kruskal.test(log(mRAP) ~ cluster6, data = Clusterpheno)
mRAP6
dunn.test(log(Clusterpheno$mRAP), Clusterpheno$cluster6, method = "BH")
```

## SvO2

Not normally distributed, unequal variance ==> log transformed

```{r, warning = FALSE, message = FALSE}
kable(cbind(table(is.na(Clusterpheno$SvO2)), prop.table(table(is.na(Clusterpheno$SvO2)))*100), col.names = c("Missing Values", "%")) %>% kable_styling(full_width = TRUE)
Clusterpheno %>% select(cluster6, SvO2) %>% filter(!is.na(SvO2)) %>% group_by(cluster6) %>% summarise(Shapiro.pvalue = shapiro.test(SvO2)$p.value, Mean = mean(SvO2), logmean = mean(log(SvO2)), Variance = var(SvO2), logvar = var(log(SvO2)), min = min(SvO2), max =  max(SvO2), median = median(SvO2), IQR = IQR(SvO2))%>% kable(.) %>% kable_styling(full_width = TRUE)

Clusterpheno %>% select(cluster6, SvO2) %>% filter(!is.na(SvO2)) %>% ggplot(aes(x= cluster6, y =SvO2)) + geom_boxplot(fill=clusterfills) + labs(x = "Cluster", y = "SvO2") + theme(panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey"))

svo26<- kruskal.test(log(SvO2) ~ cluster6, data = Clusterpheno)
svo26
dunn.test(log(Clusterpheno$SvO2), Clusterpheno$cluster6, method = "BH")
```

## PAWP

Not normally distributed, unequal variance ==> log transformed

```{r, warning = FALSE, message = FALSE}
kable(cbind(table(is.na(Clusterpheno$PAWP)), prop.table(table(is.na(Clusterpheno$PAWP)))*100), col.names = c("Missing Values", "%")) %>% kable_styling(full_width = TRUE)
Clusterpheno %>% select(cluster6, PAWP) %>% filter(!is.na(PAWP)) %>% group_by(cluster6) %>% summarise(Shapiro.pvalue = shapiro.test(PAWP)$p.value, Mean = mean(PAWP), logmean = mean(log(PAWP)), Variance = var(PAWP), logvar = var(log(PAWP)), min = min(PAWP), max =  max(PAWP), median = median(PAWP), IQR = IQR(PAWP)) %>% kable(.) %>% kable_styling(full_width = TRUE)

Clusterpheno %>% select(cluster6, PAWP) %>% filter(!is.na(PAWP)) %>% ggplot(aes(x= cluster6, y =PAWP)) + geom_boxplot(fill=clusterfills) + labs(x = "Cluster", y = "PAWP") + theme(panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey"))

pawp6<- kruskal.test(log(PAWP) ~ cluster6, data = Clusterpheno)
pawp6
dunn.test(log(Clusterpheno$PAWP), Clusterpheno$cluster6, method = "BH")
```

## Platelet Count

Not normally distributed, unequal variance ==> log transformed

```{r, warning = FALSE, message = FALSE}
kable(cbind(table(is.na(Clusterpheno$platelet_count)), prop.table(table(is.na(Clusterpheno$platelet_count)))*100), col.names = c("Missing Values", "%")) %>% kable_styling(full_width = TRUE)
Clusterpheno %>% select(cluster6, platelet_count) %>% filter(!is.na(platelet_count)) %>% group_by(cluster6) %>% summarise(Shapiro.pvalue = shapiro.test(platelet_count)$p.value, Mean = mean(platelet_count), logmean = mean(log(platelet_count)), Variance = var(platelet_count), logvar = var(log(platelet_count)), min = min(platelet_count), max =  max(platelet_count), median = median(platelet_count), IQR = IQR(platelet_count)) %>% kable(.) %>% kable_styling(full_width = TRUE)

Clusterpheno %>% select(cluster6, platelet_count) %>% filter(!is.na(platelet_count)) %>% ggplot(aes(x= cluster6, y =platelet_count)) + geom_boxplot(fill=clusterfills) + labs(x = "Cluster", y = "platelet_count") + theme(panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey"))

pc6<- kruskal.test(log(platelet_count) ~ cluster6, data = Clusterpheno)
pc6
```

## eGFR

eGFR - gfr estimated in Imperial & Cambridge, Sheffield data gfr. Sheffield units ml/min/1.73m^2 

Not normally distributed, unequal variance ==> log transformed

```{r, warning = FALSE, message = FALSE}
kable(cbind(table(is.na(Clusterpheno$eGFR)), prop.table(table(is.na(Clusterpheno$eGFR)))*100), col.names = c("Missing Values", "%")) %>% kable_styling(full_width = TRUE)
Clusterpheno %>% select(cluster6, eGFR) %>% filter(!is.na(eGFR)) %>% group_by(cluster6) %>% summarise(Shapiro.pvalue = shapiro.test(eGFR)$p.value, Mean = mean(eGFR), logmean = mean(log(eGFR)), Variance = var(eGFR), logvar = var(log(eGFR)), min = min(eGFR), max =  max(eGFR), median = median(eGFR), IQR = IQR(eGFR)) %>% kable(.) %>% kable_styling(full_width = TRUE)

Clusterpheno %>% select(cluster6, eGFR) %>% filter(!is.na(eGFR)) %>% ggplot(aes(x= cluster6, y =eGFR)) + geom_boxplot(fill=clusterfills) + labs(x = "Cluster", y = "eGFR") + theme(panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey"))

egfr6<- kruskal.test(log(eGFR) ~ cluster6, data = Clusterpheno)
egfr6
dunn.test(log(Clusterpheno$eGFR), Clusterpheno$cluster6, method = "BH")
```

## Cardiac Output

RHC fick

Not normally distributed, roughly equal variance

```{r, warning = FALSE, message = FALSE}
kable(cbind(table(is.na(Clusterpheno$CardiacOutput)), prop.table(table(is.na(Clusterpheno$CardiacOutput)))*100), col.names = c("Missing Values", "%")) %>% kable_styling(full_width = TRUE)
Clusterpheno %>% select(cluster6, CardiacOutput) %>% filter(!is.na(CardiacOutput)) %>% group_by(cluster6) %>% summarise(Shapiro.pvalue = shapiro.test(CardiacOutput)$p.value, Mean = mean(CardiacOutput), Variance = var(CardiacOutput), min = min(CardiacOutput), max =  max(CardiacOutput), median = median(CardiacOutput), IQR = IQR(CardiacOutput)) %>% kable(.) %>% kable_styling(full_width = TRUE)

Clusterpheno %>% select(cluster6, CardiacOutput) %>% filter(!is.na(CardiacOutput)) %>% ggplot(aes(x= cluster6, y =CardiacOutput)) + geom_boxplot(fill=clusterfills) + labs(x = "Cluster", y = "Cardiac Output") + theme(panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey"))

CardiacOutput6<- kruskal.test(CardiacOutput ~ cluster6, data = Clusterpheno)
CardiacOutput6
```

## Cardiac Index

CardiacIndex RHC fick

Not normally distributed, roughly equal variance

```{r, warning = FALSE, message = FALSE}
kable(cbind(table(is.na(Clusterpheno$CardiacIndex)), prop.table(table(is.na(Clusterpheno$CardiacIndex)))*100), col.names = c("Missing Values", "%")) %>% kable_styling(full_width = TRUE)
Clusterpheno %>% select(cluster6, CardiacIndex) %>% filter(!is.na(CardiacIndex)) %>% group_by(cluster6) %>% summarise(Shapiro.pvalue = shapiro.test(CardiacIndex)$p.value, Mean = mean(CardiacIndex), Variance = var(CardiacIndex), min = min(CardiacIndex), max =  max(CardiacIndex), median = median(CardiacIndex), IQR = IQR(CardiacIndex)) %>% kable(.) %>% kable_styling(full_width = TRUE)

Clusterpheno %>% select(cluster6, CardiacIndex) %>% filter(!is.na(CardiacIndex)) %>% ggplot(aes(x= cluster6, y =CardiacIndex)) + geom_boxplot(fill=clusterfills) + labs(x = "Cluster", y = "Cardiac Index") + theme(panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey"))

CardiacIndex6<- kruskal.test(CardiacIndex ~ cluster6, data = Clusterpheno)
CardiacIndex6
```

## Creatinine

Not normally distributed, unequal variance ==> log transformed

```{r, warning = FALSE, message = FALSE}
kable(cbind(table(is.na(Clusterpheno$Creatinine)), prop.table(table(is.na(Clusterpheno$Creatinine)))*100), col.names = c("Missing Values", "%")) %>% kable_styling(full_width = TRUE)
Clusterpheno %>% select(cluster6, Creatinine) %>% filter(!is.na(Creatinine)) %>% group_by(cluster6) %>% summarise(Shapiro.pvalue = shapiro.test(Creatinine)$p.value, Mean = mean(Creatinine), logmean = mean(log(Creatinine)), Variance = var(Creatinine), logvar = var(log(Creatinine)), min = min(Creatinine), max =  max(Creatinine), median = median(Creatinine), IQR = IQR(Creatinine)) %>% kable(.) %>% kable_styling(full_width = TRUE)

Clusterpheno %>% select(cluster6, Creatinine) %>% filter(!is.na(Creatinine)) %>% ggplot(aes(x= cluster6, y =Creatinine)) + geom_boxplot(fill=clusterfills) + labs(x = "Cluster", y = "Creatinine") + theme(panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey"))

Creat6<- kruskal.test(log(Creatinine) ~ cluster6, data = Clusterpheno)
Creat6
dunn.test(log(Clusterpheno$Creatinine), Clusterpheno$cluster6, method = "BH")
```

## TLco predicted

Not normally distributed, unequal variance ==> log transformed

```{r, warning = FALSE, message = FALSE}
kable(cbind(table(is.na(Clusterpheno$tlco_predicted )), prop.table(table(is.na(Clusterpheno$tlco_predicted)))*100), col.names = c("Missing Values", "%")) %>% kable_styling(full_width = TRUE)
Clusterpheno %>% select(cluster6, tlco_predicted) %>% filter(!is.na(tlco_predicted)) %>% group_by(cluster6) %>% summarise(Shapiro.pvalue = shapiro.test(tlco_predicted)$p.value, Mean = mean(tlco_predicted), logmean = mean(log(tlco_predicted)), Variance = var(tlco_predicted), logvar = var(log(tlco_predicted)), min = min(tlco_predicted), max =  max(tlco_predicted), median = median(tlco_predicted), IQR = IQR(tlco_predicted)) %>% kable(.) %>% kable_styling(full_width = TRUE)

Clusterpheno %>% select(cluster6, tlco_predicted) %>% filter(!is.na(tlco_predicted)) %>% ggplot(aes(x= cluster6, y =tlco_predicted)) + geom_boxplot(fill=clusterfills) + labs(x = "Cluster", y = "tlco_predicted") + theme(panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey"))

tlco6<- kruskal.test(log(tlco_predicted) ~ cluster6, data = Clusterpheno)
tlco6
dunn.test(log(Clusterpheno$tlco_predicted), Clusterpheno$cluster6, method = "BH")
```



## Comorbidity: COPD

```{r, warning = FALSE, message = FALSE}
kable(addmargins(table(is.na(Clusterpheno$copd ))), col.names = c("Missing Values", ""))

kable(addmargins(table(Clusterpheno$copd, Clusterpheno$cluster6))) %>% kable_styling(full_width = TRUE)
copd6<- fisher.test(table(Clusterpheno$copd, Clusterpheno$cluster6), conf.level = 0.95, simulate.p.value=TRUE)

copd6

as.data.frame(table(Clusterpheno$cluster6, Clusterpheno$copd)) %>% ggplot(aes(x = Var1, y = Freq, fill = Var2)) + geom_bar(stat = "identity") + scale_fill_manual(values = c("cornflowerblue","midnight blue")) + ggtitle("COPD") + labs(x = "Cluster", y = "") + guides(fill=guide_legend(title=NULL)) + theme(panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey"))
```

## Comorbidity: Sleep Apnoea


```{r, warning = FALSE, message = FALSE}
kable(addmargins(table(is.na(Clusterpheno$SleepApnoea ))), col.names = c("Missing Values", ""))

kable(addmargins(table(Clusterpheno$SleepApnoea, Clusterpheno$cluster6))) %>% kable_styling(full_width = TRUE)
SleepApnoea6<- fisher.test(table(Clusterpheno$SleepApnoea, Clusterpheno$cluster6), conf.level = 0.95)
SleepApnoea6

as.data.frame(table(Clusterpheno$cluster6, Clusterpheno$SleepApnoea)) %>% ggplot(aes(x = Var1, y = Freq, fill = Var2)) + geom_bar(stat = "identity") + scale_fill_manual(values = c("cornflowerblue","midnight blue")) + ggtitle("Sleep Apnoea") + labs(x = "Cluster", y = "") + guides(fill=guide_legend(title=NULL)) + theme(panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey"))
```

## Comorbidity: Atrial Fibrillation


```{r, warning = FALSE, message = FALSE}
kable(table(is.na(Clusterpheno$AtrialFibrillation )), col.names = c("Missing Values", ""))

kable(addmargins(table(Clusterpheno$AtrialFibrillation, Clusterpheno$cluster6))) %>% kable_styling(full_width = TRUE)
AtrialFibrillation6<- chisq.test(table(Clusterpheno$AtrialFibrillation, Clusterpheno$cluster6))
AtrialFibrillation6

kable(chisq.posthoc.test(table(Clusterpheno$AtrialFibrillation, Clusterpheno$cluster6), method = "BH"))  %>% kable_styling(full_width = TRUE)

as.data.frame(table(Clusterpheno$cluster6, Clusterpheno$AtrialFibrillation)) %>% ggplot(aes(x = Var1, y = Freq, fill = Var2)) + geom_bar(stat = "identity") + scale_fill_manual(values = c("cornflowerblue","midnight blue")) + ggtitle("Atrial Fibrillation") + labs(x = "Cluster", y = "") + guides(fill=guide_legend(title=NULL)) + theme(panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey"))
```

## Comorbidity: Type 2 Diabetes mellitus


```{r, warning = FALSE, message = FALSE}
kable(addmargins(table(is.na(Clusterpheno$t2dm ))), col.names = c("Missing Values", ""))

kable(addmargins(table(Clusterpheno$t2dm, Clusterpheno$cluster6))) %>% kable_styling(full_width = TRUE)
Diabetes6<- fisher.test(table(Clusterpheno$t2dm, Clusterpheno$cluster6), conf.level = 0.95, simulate.p.value=TRUE)
Diabetes6

as.data.frame(table(Clusterpheno$cluster6, Clusterpheno$t2dm)) %>% ggplot(aes(x = Var1, y = Freq, fill = Var2)) + geom_bar(stat = "identity") + scale_fill_manual(values = c("cornflowerblue","midnight blue")) + ggtitle("Type 2 Diabetes mellitus") + labs(x = "Cluster", y = "") + guides(fill=guide_legend(title=NULL)) + theme(panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey"))
```


## Comorbidity: Thyroid disease


```{r, warning = FALSE, message = FALSE}
kable(addmargins(table(is.na(Clusterpheno$thyroid_disease ))), col.names = c("Missing Values", ""))

kable(addmargins(table(Clusterpheno$thyroid_disease, Clusterpheno$cluster6))) %>% kable_styling(full_width = TRUE)
thyroid_disease6<- fisher.test(table(Clusterpheno$thyroid_disease, Clusterpheno$cluster6), conf.level = 0.95)
thyroid_disease6

as.data.frame(table(Clusterpheno$cluster6, Clusterpheno$thyroid_disease)) %>% ggplot(aes(x = Var1, y = Freq, fill = Var2)) + geom_bar(stat = "identity") + scale_fill_manual(values = c("cornflowerblue","midnight blue")) + ggtitle("Thyroid Disease") + labs(x = "Cluster", y = "") + guides(fill=guide_legend(title=NULL)) + theme(panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey"))
```

## Comorbidity: CTD

Not including SSc

```{r, warning = FALSE, message = FALSE}
kable(addmargins(table(is.na(Clusterpheno$OtherCTD ))), col.names = c("Missing Values", ""))

kable(addmargins(table(Clusterpheno$OtherCTD, Clusterpheno$cluster6))) %>% kable_styling(full_width = TRUE)
CTD6<- fisher.test(table(Clusterpheno$OtherCTD, Clusterpheno$cluster6), conf.level = 0.95, simulate.p.value = TRUE)
CTD6

as.data.frame(table(Clusterpheno$cluster6, Clusterpheno$OtherCTD)) %>% ggplot(aes(x = Var1, y = Freq, fill = Var2)) + geom_bar(stat = "identity") + scale_fill_manual(values = c("cornflowerblue","midnight blue")) + ggtitle("Other CTD") + labs(x = "Cluster", y = "") + guides(fill=guide_legend(title=NULL)) + theme(panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey"))
```

## Comorbidity: SSc


```{r, warning = FALSE, message = FALSE}
kable(addmargins(table(is.na(Clusterpheno$ssc ))), col.names = c("Missing Values", ""))

kable(addmargins(table(Clusterpheno$ssc, Clusterpheno$cluster6))) %>% kable_styling(full_width = TRUE)
scleroderma6<- fisher.test(table(Clusterpheno$ssc, Clusterpheno$cluster6), conf.level = 0.95, simulate.p.value = TRUE)
scleroderma6

as.data.frame(table(Clusterpheno$cluster6, Clusterpheno$ssc)) %>% ggplot(aes(x = Var1, y = Freq, fill = Var2)) + geom_bar(stat = "identity") + scale_fill_manual(values = c("cornflowerblue","midnight blue")) + ggtitle("SSc") + labs(x = "Cluster", y = "") + guides(fill=guide_legend(title=NULL)) + theme(panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey"))
```

## Comorbidity: Ischaemic heart disease


```{r, warning = FALSE, message = FALSE}
kable(addmargins(table(is.na(Clusterpheno$ihd ))), col.names = c("Missing Values", ""))

Clusterpheno$ihd <- replace_na(Clusterpheno$ihd, "no")
kable(addmargins(table(Clusterpheno$ihd, Clusterpheno$cluster6))) %>% kable_styling(full_width = TRUE)
ihd6<- fisher.test(table(Clusterpheno$ihd, Clusterpheno$cluster6), conf.level = 0.95, simulate.p.value = TRUE)
ihd6

kable(chisq.posthoc.test(table(Clusterpheno$ihd, Clusterpheno$cluster6), method = "BH"))  %>% kable_styling(full_width = TRUE)

as.data.frame(table(Clusterpheno$cluster6, Clusterpheno$ihd)) %>% ggplot(aes(x = Var1, y = Freq, fill = Var2)) + geom_bar(stat = "identity") + scale_fill_manual(values = c("cornflowerblue","midnight blue")) + ggtitle("Ischaemic heart disease") + labs(x = "Cluster", y = "") + guides(fill=guide_legend(title=NULL)) + theme(panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey"))
```





<a id="Summary"></a> 

# Summary

```{r}
clinsums <- as.data.frame(rbind(c("sex", sex6$p.value, ""), c("FC", fc6$p.value, ""), c("Site", site6$p.value, "B,C,D,E"), c("BMI", bmi6$p.value, ""), c("Age", age6$p.value, ""), c("NTproBNP", ntprobnp6$p.value, "A/C, B/C, C/D, C/F, D/E"), c("Uric Acid", ua6$p.value, "C/D"), c("FVCP", 0.394, ""), c("REVEAL risk group", Reveal6$p.value, ""), c("Treatment", treatment6$p.value, "C & D"), c("6MWD", smwd6$p.value, "A/E, B/E, D/E, E/F"), c("ISWT", iswt6$p.value, ""), c("BPDIA",bpdia6$p.value, "" ), c("BPSYS", bpsys6$p.value, ""), c("PVR", pvr6$p.value, "A/C, A/E, E/F"), c("mPAP", mpap6$p.value, "A/B, A/C, A/E, E/F"), c("sPAP", spap6$p.value, "A/B, A/C, A/D, A/E, E/F"), c("dPAP", dpap6$p.value, "A/B, A/C, A/E"), c("mRAP", mRAP6$p.value, "A/C, B/C, C/D, C/E, C/F"), c("SvO2", svo26$p.value, "A/C, A/E, B/C, B/E, C/D, C/F, D/E, E/F"), c("PAWP", pawp6$p.value, ""), c("Platelet Count", pc6$p.value, ""), c("eGFR", egfr6$p.value, "B/C, B/D, B/E, D/F"), c("Cardiac Output", CardiacOutput6$p.value, ""), c("Cardiac Index", CardiacIndex6$p.value, ""), c("Creatinine", Creat6$p.value, "C/D"), c("TLCO predicted", tlco6$p.value, ""), c("Comorbidity: COPD", copd6$p.value, ""), c("Comorbidity: Sleep Apnea", SleepApnoea6$p.value, ""), c("Comorbidity: Atrial Fibrillation", AtrialFibrillation6$p.value, ""), c("Comorbidity: Diabetes", Diabetes6$p.value, ""), c("Comorbidity: Thyroid Disease", thyroid_disease6$p.value, ""), c("Comorbidity: CTD", CTD6$p.value, ""), c("Comorbidity: Scleroderma", scleroderma6$p.value, ""), c("Comorbidity: Ischaemic heart disease", ihd6$p.value, "" )))
clinsums <- clinsums %>% mutate(V2 = as.numeric(V2)) %>% mutate(V2 = signif(V2, 3)) %>% rename(`Clinical Parameter` = V1, Pvalue = V2, `Significant post-hoc tests` = V3) %>% mutate(`Adjusted Pvalue` = signif(p.adjust(Pvalue, "BH"), 3)) %>% mutate(Significant = ifelse(`Adjusted Pvalue` <= 0.05, "*", "")) %>% select(`Clinical Parameter`, Pvalue, `Adjusted Pvalue`, Significant, `Significant post-hoc tests`) %>% arrange(`Adjusted Pvalue`)

kable(clinsums) %>% kable_styling(full_width = TRUE)
```



```{r, eval = F}
# MicroRNA signatures
#Where signatures have been derrived for comparisons between groups, how much overlap is in these. Max number per signature = 9. Grey boxes indicate 0 miRNA overlap. NTproBNP not included.

library(readxl)
library(pheatmap)
summarytable <- read_excel("/Volumes/GoogleDrive/My Drive/Actelion/UsingUpdatedDANA/miRNA clustering/Run 5 - MIREXS Phase DI (PH1, PH2, PH3) on Chris data/Run5Summary.xlsx", 
    sheet = "Summary")
Overlap <- read_excel("/Volumes/GoogleDrive/My Drive/Actelion/UsingUpdatedDANA/miRNA clustering/Run 5 - MIREXS Phase DI (PH1, PH2, PH3) on Chris data/Run5Summary.xlsx", 
    sheet = "Overlap")

Overlap %>% 
  as.data.frame() %>%
  pivot_longer(-c(X), names_to = "Comparison", values_to = "counts") %>% 
  ggplot(aes(x=Comparison, y=X, fill=counts)) + 
  geom_tile() + theme_minimal()+ theme(axis.text.x = element_text(angle = 90, vjust = 1, size = 12, hjust = 1)) + scale_color_brewer()

#+ scale_fill_gradient2(low = "red",  mid = "white",  high = "blue",  midpoint = 0,  space = "Lab",  na.value = "grey50",  guide = "colourbar",  aesthetics = "fill")
```

# LASSO miRNA cluster signatures

```{r, message = FALSE}
LASSOmiRs <- read_csv("LASSO/LASSO.miR.Caret.coefficients.csv")
LASSOmiRs$miR <- gsub(x = LASSOmiRs$miR, "\\.", "-")
library(viridis)
allcoefslong <- LASSOmiRs %>% pivot_longer(-c(miR), names_to = "Cluster", values_to = "Coefficient") %>% filter(Coefficient != 0) %>% arrange(desc(abs(Coefficient)))  %>%  mutate(miR=factor(miR, levels = unique(miR)))

ggplot(allcoefslong, aes(x = miR, y = Coefficient, fill = Cluster)) + geom_col() + theme_minimal()+ theme(axis.text.x = element_text(angle = 90, vjust = 1, size = 12, hjust = 1)) + scale_fill_viridis(discrete=TRUE)

lassoA <- allcoefslong %>% filter(Cluster == "clusterA") 
lassoB <- allcoefslong %>% filter(Cluster == "clusterB") 
lassoC <- allcoefslong %>% filter(Cluster == "clusterC") 
lassoD <- allcoefslong %>% filter(Cluster == "clusterD") 
lassoE <- allcoefslong %>% filter(Cluster == "clusterE") 
lassoF <- allcoefslong %>% filter(Cluster == "clusterF") 

```

## Cluster A signature

MicroRNAs in the cluster A signature. Cluster A values in pink. Training & Interim sets

```{r}
select(Clusterpheno, c("cluster6", lassoA$miR)) %>% as.data.frame() %>%  pivot_longer(-c(cluster6), names_to = "miR", values_to = "Expression") %>% ggplot(aes(x = miR, y = Expression)) + geom_boxplot(aes(fill = cluster6)) + theme(text = element_text(size=12), axis.text.x=element_text(angle = 45,vjust=0.5, size = 12),axis.text.y=element_text(size = 12),legend.position = "none", panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey")) + labs(x="", size = 12) +scale_fill_manual(values=c("deeppink","cornflowerblue", "cornflowerblue", "cornflowerblue", "cornflowerblue", "cornflowerblue")) + ggtitle("Cluster A")


```

## Cluster B signature

MicroRNAs in the cluster B signature. Cluster B values in red. Training & Interim sets

```{r}
select(Clusterpheno, c("cluster6", lassoB$miR))  %>% as.data.frame() %>%  pivot_longer(-c(cluster6), names_to = "miR", values_to = "Expression") %>% ggplot(aes(x = miR, y = Expression)) + geom_boxplot(aes(fill = cluster6)) + theme(text = element_text(size=12), axis.text.x=element_text(angle = 45,vjust=0.5, size = 12),axis.text.y=element_text(size = 12),legend.position = "none", panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey")) + labs(x="", size = 12) +scale_fill_manual(values=c("cornflowerblue","deeppink", "cornflowerblue", "cornflowerblue", "cornflowerblue", "cornflowerblue")) + ggtitle("Cluster B")
```

## Cluster C signature

MicroRNAs in the cluster C signature. Cluster C values in red. Training & Interim sets

```{r}
select(Clusterpheno, c("cluster6", lassoC$miR)) %>% as.data.frame() %>%  pivot_longer(-c(cluster6), names_to = "miR", values_to = "Expression") %>% ggplot(aes(x = miR, y = Expression)) + geom_boxplot(aes(fill = cluster6)) + theme(text = element_text(size=12), axis.text.x=element_text(angle = 60,vjust=0.5, size = 12),axis.text.y=element_text(size = 12),legend.position = "none", panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey"))+ labs(x="", size = 12) +scale_fill_manual(values=c("cornflowerblue","cornflowerblue", "deeppink", "cornflowerblue", "cornflowerblue", "cornflowerblue")) + ggtitle("Cluster C")
```

## Cluster D signature

MicroRNAs in the cluster D signature. Cluster D values in red. Training & Interim sets

```{r}
select(Clusterpheno, c("cluster6", lassoD$miR)) %>% as.data.frame() %>%  pivot_longer(-c(cluster6), names_to = "miR", values_to = "Expression") %>% ggplot(aes(x = miR, y = Expression)) + geom_boxplot(aes(fill = cluster6)) + theme(text = element_text(size=12), axis.text.x=element_text(angle = 60,vjust=0.5, size = 12),axis.text.y=element_text(size = 12),legend.position = "none", panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey")) + labs(x="", size = 12) +scale_fill_manual(values=c("cornflowerblue","cornflowerblue", "cornflowerblue", "deeppink", "cornflowerblue", "cornflowerblue")) + ggtitle("Cluster D")
```

## Cluster E signature

MicroRNAs in the cluster E signature. Cluster E values in red. Training & Interim sets

```{r}
select(Clusterpheno, c("cluster6", lassoE$miR))  %>% as.data.frame() %>%  pivot_longer(-c(cluster6), names_to = "miR", values_to = "Expression") %>% ggplot(aes(x = miR, y = Expression)) + geom_boxplot(aes(fill = cluster6)) + theme(text = element_text(size=12), axis.text.x=element_text(angle = 45,vjust=0.5, size = 12),axis.text.y=element_text(size = 12),legend.position = "none", panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey")) + labs(x="", size = 12) +scale_fill_manual(values=c("cornflowerblue","cornflowerblue", "cornflowerblue", "cornflowerblue", "deeppink", "cornflowerblue")) + ggtitle("Cluster E")
```

## Cluster F signature

MicroRNAs in the cluster F signature. Cluster F values in red. Training & Interim sets

```{r}
select(Clusterpheno, c("cluster6", lassoF$miR)) %>% as.data.frame() %>%  pivot_longer(-c(cluster6), names_to = "miR", values_to = "Expression") %>% ggplot(aes(x = miR, y = Expression)) + geom_boxplot(aes(fill = cluster6)) + theme(text = element_text(size=12), axis.text.x=element_text(angle = 60,vjust=0.5, size = 12),axis.text.y=element_text(size = 12),legend.position = "none", panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey")) + labs(x="", size = 12) +scale_fill_manual(values=c("cornflowerblue","cornflowerblue", "cornflowerblue", "cornflowerblue", "cornflowerblue", "deeppink")) + ggtitle("Cluster F")
```

# LASSO coefficients 


```{r}
allcoefslong %>% 
  ggplot(aes(x=Cluster, y=miR, fill=Coefficient)) + 
  geom_tile() + theme_minimal()+ theme(axis.text.x = element_text(angle = 90, vjust = 1, size = 12, hjust = 1)) + 
  scale_fill_gradientn(colours = c("midnight blue", "white", "dark red"), limits = c(-1.1,1.1))

```


```{r}
ggplot(lassoA, aes(x = reorder(miR, Coefficient), y = Coefficient, fill = Coefficient > 0)) + geom_col() + theme(text = element_text(size=12), axis.text.x=element_text(angle = 90,vjust=0.5, size = 12),axis.text.y=element_text(size = 12), legend.position = "none", panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey")) + labs(x="", size = 12) + ggtitle("Cluster A") + scale_fill_manual(values = c("cornflowerblue", "deeppink"))
```

```{r}
ggplot(lassoB, aes(x = reorder(miR, Coefficient), y = Coefficient, fill = Coefficient > 0)) + geom_col() + theme(text = element_text(size=12), axis.text.x=element_text(angle = 90,vjust=0.5, size = 12),axis.text.y=element_text(size = 12), legend.position = "none", panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey")) + labs(x="", size = 12) + ggtitle("Cluster B") + scale_fill_manual(values = c("cornflowerblue", "deeppink"))
```

```{r}
ggplot(lassoC, aes(x = reorder(miR, Coefficient), y = Coefficient, fill = Coefficient > 0)) + geom_col() + theme(text = element_text(size=12), axis.text.x=element_text(angle = 90,vjust=0.5, size = 12),axis.text.y=element_text(size = 12), legend.position = "none", panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey")) + labs(x="", size = 12) + ggtitle("Cluster C") + scale_fill_manual(values = c("cornflowerblue", "deeppink"))
```

```{r}
ggplot(lassoD, aes(x = reorder(miR, Coefficient), y = Coefficient, fill = Coefficient > 0)) + geom_col() + theme(text = element_text(size=12), axis.text.x=element_text(angle = 90,vjust=0.5, size = 12),axis.text.y=element_text(size = 12), legend.position = "none", panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey")) + labs(x="", size = 12) + ggtitle("Cluster D") + scale_fill_manual(values = c("cornflowerblue", "deeppink"))
```

```{r}
ggplot(lassoE, aes(x = reorder(miR, Coefficient), y = Coefficient, fill = Coefficient > 0)) + geom_col() + theme(text = element_text(size=12), axis.text.x=element_text(angle = 90,vjust=0.5, size = 12),axis.text.y=element_text(size = 12), legend.position = "none", panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey")) + labs(x="", size = 12) + ggtitle("Cluster E") + scale_fill_manual(values = c("cornflowerblue", "deeppink"))
```

```{r}
ggplot(lassoF, aes(x = reorder(miR, Coefficient), y = Coefficient, fill = Coefficient > 0)) + geom_col() + theme(text = element_text(size=12), axis.text.x=element_text(angle = 90,vjust=0.5, size = 12),axis.text.y=element_text(size = 12), legend.position = "none", panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey")) + labs(x="", size = 12) + ggtitle("Cluster F") + scale_fill_manual(values = c("cornflowerblue", "deeppink"))
```

# Survival

SampleID - sampleID to match to cluster membership

DaysToEvent - number of days between sample (not necessarily diagnosis) and census date OR date of death. NB Imperial provided 'Days after diagnosis: sample date', 'days after diagnosis: Death', 'Status Alive/Dead', 'Days after diagnosis: Last follow up'. NOT time to census - some 'Days after diagnosis: Last follow up' are negative. Where this leads to a negative DaysToEvent, numbers set to 0.

Survivor: survival status of patient at census date (1 = dead, 0 = alive)

12 Patients who have undergone transplant were removed before analysis. (2 Cambridge, 2 Imperial, 8 Sheffield)

```{r}
library(survival)
library(survminer)
SurviveData <- read_excel("../Survival.xlsx", sheet = "All")
#SampleID - sampleID to match to cluster membership
#DaysToEvent - number of days between sample (not necessarily diagnosis) and census date OR date of death. NB Imperial provided 'Days after diagnosis: sample date', 'days after diagnosis: Death', 'Status Alive/Dead', 'Days after diagnosis: Last follow up'. NOT time to census - some 'Days after diagnosis: Last follow up' are negative. Where this leads to a negative DaysToEvent, numbers set to 0.
#Survivor: survival status of patient at census date (1 = dead, 0 = alive)

#SurviveData <- SurviveData %>% filter(Transplant == 'n')
#Clusterpheno <- read_csv("Clusterpheno.csv")

ForSurvival <- left_join(Clusterpheno, SurviveData, by = "SampleID")
Bases<- survival::Surv(time = ForSurvival$DaysToEvent, event = ForSurvival$Survivor)
Bases1<- survival::Surv(time = ForSurvival$DaysToEvent1yr, event = ForSurvival$Survivor1yr)
Bases3<- survival::Surv(time = ForSurvival$DaysToEvent3yr, event = ForSurvival$Survivor3yr)
Bases5<- survival::Surv(time = ForSurvival$DaysToEvent5yr, event = ForSurvival$Survivor5yr)
Bases10<- survival::Surv(time = ForSurvival$DaysToEvent10yr, event = ForSurvival$Survivor10yr)

survminer::ggsurvplot(
    fit = survival::survfit(Bases ~ 1, data = ForSurvival), 
    xlab = "Years", 
    ylab = "Overall survival probability",
    legend.title = "", risk.table = TRUE, xscale = "d_y")

summary(survival::survfit(Bases ~ 1, data = ForSurvival), times = 365.25)
```


Median survival time: 

```{r}
survival::survfit(Bases ~ 1, data = ForSurvival)
```

## Age and sex

```{r}
survminer::ggsurvplot(
    fit = survival::survfit(Bases ~ sex, data = ForSurvival), 
    xlab = "Days", 
    ylab = "Overall survival probability",
    legend.title = "",
    palette=c("deeppink", "cornflowerblue"),
    conf.int = TRUE, risk.table = TRUE)

broom::tidy(coxph(Bases~sex+age, data = ForSurvival), exp = TRUE) %>% kable(., caption = "Cox regression") %>% kable_styling(full_width = TRUE)

library(gtsummary)
coxph(Bases~sex+age, data = ForSurvival) %>% gtsummary::tbl_regression(exp = TRUE)
```

```{r}
ForSurvival %>% group_by(Survivor10yr) %>% summarise (mean = mean(DaysToEvent10yr), median = median(DaysToEvent10yr)) %>% kable(., caption = "Days to Event") %>% kable_styling(full_width = TRUE)

kable(cbind(table(ForSurvival$Survivor10yr), prop.table(table(ForSurvival$Survivor10yr))*100), col.names = c("Survival Status (n)", "%")) %>% kable_styling(full_width = TRUE)

ForSurvival %>% summarise(n = n(), mean = mean(DaysToEvent10yr), median = median(DaysToEvent10yr)) %>% kable() %>% kable_styling(full_width = TRUE)

ForSurvival %>% mutate(`Age Bracket` = cut(age, breaks=c(0, 65, Inf), labels = c("below median age", "above median age"))) %>%   group_by(`Age Bracket`, Survivor10yr) %>%
  summarise(n = n()) %>%   mutate(`%` = (n / sum(n))*100) %>% kable() %>% kable_styling(full_width = TRUE)

kable(cbind(table(ForSurvival$sex, ForSurvival$Survivor10yr), prop.table(table(ForSurvival$sex, ForSurvival$Survivor10yr),1)*100), col.names = c("Survivor (n)", "Deceased (n)", "Survivor (%)", "Deceased (%)"), caption = "Survival status by sex") %>% kable_styling(full_width = TRUE)

```


## Cluster survival

### All time available

```{r}
survminer::ggsurvplot(
    fit = survival::survfit(Bases ~ cluster6, data = ForSurvival), 
    xlab = "Years", 
    ylab = "Overall survival probability",
    legend.title = "",
    palette=c("deeppink","dark red", "darksalmon", "yellow", "midnight blue", "cornflowerblue"),
    risk.table = TRUE,  risk.table.height = 0.25, xscale = "d_y", break.time.by=365.25*2.5)


broom::tidy(coxph(Bases~sex+age+cluster6, data = ForSurvival), exp = TRUE) %>% kable(., caption = "Cox regression: ~ sex + age + 6 clusters") %>% kable_styling(full_width = TRUE)
coxph(Bases~sex+age+cluster6, data = ForSurvival) %>% gtsummary::tbl_regression(exp = TRUE)

broom::tidy(coxph(Bases~cluster6, data = ForSurvival), exp = TRUE) %>% kable(., caption = "Cox regression: ~ 6 clusters") %>% kable_styling(full_width = TRUE)
coxph(Bases~cluster6, data = ForSurvival) %>% gtsummary::tbl_regression(exp = TRUE)

ggforest(coxph(Bases~sex+age+cluster6, data = ForSurvival))
```

### 10 year survival

```{r}
survminer::ggsurvplot(
    fit = survival::survfit(Bases10 ~ cluster6, data = ForSurvival), 
    xlab = "Years", 
    ylab = "Overall survival probability",
    legend.title = "",
    palette=c("deeppink","dark red", "darksalmon", "yellow", "midnight blue", "cornflowerblue"),
    risk.table = TRUE,  risk.table.height = 0.25, xscale = "d_y", break.time.by=365.25*2.5)


broom::tidy(coxph(Bases10~sex+age+cluster6, data = ForSurvival), exp = TRUE) %>% kable(., caption = "Cox regression: ~ sex + age + 6 clusters") %>% kable_styling(full_width = TRUE)
coxph(Bases10~sex+age+cluster6, data = ForSurvival) %>% gtsummary::tbl_regression(exp = TRUE)

broom::tidy(coxph(Bases10~cluster6, data = ForSurvival), exp = TRUE) %>% kable(., caption = "Cox regression: ~ 6 clusters") %>% kable_styling(full_width = TRUE)
coxph(Bases10~cluster6, data = ForSurvival) %>% gtsummary::tbl_regression(exp = TRUE)

ggforest(coxph(Bases10~sex+age+cluster6, data = ForSurvival))
```

### 5 year survival

```{r}
survminer::ggsurvplot(
    fit = survival::survfit(Bases5 ~ cluster6, data = ForSurvival), 
    xlab = "Years", 
    ylab = "Overall survival probability",
    legend.title = "",
    palette=c("deeppink","dark red", "darksalmon", "yellow", "midnight blue", "cornflowerblue"),
    risk.table = TRUE,  risk.table.height = 0.25, xscale = "d_y", break.time.by=365.25*2.5)


broom::tidy(coxph(Bases5~sex+age+cluster6, data = ForSurvival), exp = TRUE) %>% kable(., caption = "Cox regression: ~ sex + age + 6 clusters") %>% kable_styling(full_width = TRUE)
coxph(Bases5~sex+age+cluster6, data = ForSurvival) %>% gtsummary::tbl_regression(exp = TRUE)

broom::tidy(coxph(Bases5~cluster6, data = ForSurvival), exp = TRUE) %>% kable(., caption = "Cox regression: ~ 6 clusters") %>% kable_styling(full_width = TRUE)
coxph(Bases5~cluster6, data = ForSurvival) %>% gtsummary::tbl_regression(exp = TRUE)

ggforest(coxph(Bases5~sex+age+cluster6, data = ForSurvival))
```

### 3 year survival

```{r}
survminer::ggsurvplot(
    fit = survival::survfit(Bases3 ~ cluster6, data = ForSurvival), 
    xlab = "Years", 
    ylab = "Overall survival probability",
    legend.title = "",
    palette=c("deeppink","dark red", "darksalmon", "yellow", "midnight blue", "cornflowerblue"),
    risk.table = TRUE,  risk.table.height = 0.25, xscale = "d_y", break.time.by=365.25*2.5)


broom::tidy(coxph(Bases3~sex+age+cluster6, data = ForSurvival), exp = TRUE) %>% kable(., caption = "Cox regression: ~ sex + age + 6 clusters") %>% kable_styling(full_width = TRUE)
coxph(Bases3~sex+age+cluster6, data = ForSurvival) %>% gtsummary::tbl_regression(exp = TRUE)

broom::tidy(coxph(Bases3~cluster6, data = ForSurvival), exp = TRUE) %>% kable(., caption = "Cox regression: ~ 6 clusters") %>% kable_styling(full_width = TRUE)
coxph(Bases3~cluster6, data = ForSurvival) %>% gtsummary::tbl_regression(exp = TRUE)

ggforest(coxph(Bases3~sex+age+cluster6, data = ForSurvival))
```

### 1 year survival

```{r}
survminer::ggsurvplot(
    fit = survival::survfit(Bases1 ~ cluster6, data = ForSurvival), 
    xlab = "Years", 
    ylab = "Overall survival probability",
    legend.title = "",
    palette=c("deeppink","dark red", "darksalmon", "yellow", "midnight blue", "cornflowerblue"))


broom::tidy(coxph(Bases1~sex+age+cluster6, data = ForSurvival), exp = TRUE) %>% kable(., caption = "Cox regression: ~ sex + age + 6 clusters") %>% kable_styling(full_width = TRUE)
coxph(Bases1~sex+age+cluster6, data = ForSurvival) %>% gtsummary::tbl_regression(exp = TRUE)

broom::tidy(coxph(Bases1~cluster6, data = ForSurvival), exp = TRUE) %>% kable(., caption = "Cox regression: ~ 6 clusters") %>% kable_styling(full_width = TRUE)
coxph(Bases1~cluster6, data = ForSurvival) %>% gtsummary::tbl_regression(exp = TRUE)

ggforest(coxph(Bases1~sex+age+cluster6, data = ForSurvival))
```

## WHO clinical classification group survival

```{r}
ForSurvivalA <- ForSurvival %>% filter(cluster6 == "A")
BasesA<- survival::Surv(time = ForSurvivalA$DaysToEvent10yr, event = ForSurvivalA$Survivor10yr)
ForSurvivalB <- ForSurvival %>% filter(cluster6 == "B")
BasesB<- survival::Surv(time = ForSurvivalB$DaysToEvent10yr, event = ForSurvivalB$Survivor10yr)
ForSurvivalC <- ForSurvival %>% filter(cluster6 == "C")
BasesC<- survival::Surv(time = ForSurvivalC$DaysToEvent10yr, event = ForSurvivalC$Survivor10yr)
ForSurvivalD <- ForSurvival %>% filter(cluster6 == "D")
BasesD<- survival::Surv(time = ForSurvivalD$DaysToEvent10yr, event = ForSurvivalD$Survivor10yr)
ForSurvivalE <- ForSurvival %>% filter(cluster6 == "E")
BasesE<- survival::Surv(time = ForSurvivalE$DaysToEvent10yr, event = ForSurvivalE$Survivor10yr)
ForSurvivalF <- ForSurvival %>% filter(cluster6 == "F")
BasesF<- survival::Surv(time = ForSurvivalF$DaysToEvent10yr, event = ForSurvivalF$Survivor10yr)

survminer::ggsurvplot(
    fit = survival::survfit(Bases10 ~ dana, data = ForSurvival), 
    xlab = "Years", 
    ylab = "Overall survival probability",
    legend.title = "",
    palette=c("deeppink","dark red", "darksalmon", "yellow", "midnight blue", "cornflowerblue"),
    risk.table = TRUE,  risk.table.height = 0.25, xscale = "d_y", break.time.by=365.25*2.5)

```

```{r}
survminer::ggsurvplot(
    fit = survival::survfit(Bases10 ~ dana1, data = ForSurvival), 
    xlab = "Years", 
    ylab = "Overall survival probability",
    legend.title = "",
    palette=c("deeppink", "midnight blue", "cornflowerblue"),
    risk.table = TRUE,  risk.table.height = 0.25, xscale = "d_y", break.time.by=365.25*2.5)

```


```{r}
survminer::ggsurvplot(
    fit = survival::survfit(BasesA ~ dana1, data = ForSurvivalA), 
    xlab = "Years", 
    ylab = "Overall survival probability",
    legend.title = "",
    palette=c("deeppink", "midnight blue", "cornflowerblue"),
    risk.table = TRUE,  risk.table.height = 0.25, xscale = "d_y", break.time.by=365.25*2.5) + ggtitle("Cluster A")

survminer::ggsurvplot(
    fit = survival::survfit(BasesB ~ dana1, data = ForSurvivalB), 
    xlab = "Years", 
    ylab = "Overall survival probability",
    legend.title = "",
    palette=c("deeppink", "midnight blue", "cornflowerblue"),
    risk.table = TRUE,  risk.table.height = 0.25, xscale = "d_y", break.time.by=365.25*2.5) + ggtitle("Cluster B")

survminer::ggsurvplot(
    fit = survival::survfit(BasesC ~ dana1, data = ForSurvivalC), 
    xlab = "Years", 
    ylab = "Overall survival probability",
    legend.title = "",
    palette=c("deeppink", "midnight blue", "cornflowerblue"),
    risk.table = TRUE,  risk.table.height = 0.25, xscale = "d_y", break.time.by=365.25*2.5) + ggtitle("Cluster C")

survminer::ggsurvplot(
    fit = survival::survfit(BasesD ~ dana1, data = ForSurvivalD), 
    xlab = "Years", 
    ylab = "Overall survival probability",
    legend.title = "",
    palette=c("deeppink", "midnight blue", "cornflowerblue"),
    risk.table = TRUE,  risk.table.height = 0.25, xscale = "d_y", break.time.by=365.25*2.5) + ggtitle("Cluster D")

survminer::ggsurvplot(
    fit = survival::survfit(BasesE ~ dana1, data = ForSurvivalE), 
    xlab = "Years", 
    ylab = "Overall survival probability",
    legend.title = "",
    palette=c("deeppink", "midnight blue", "cornflowerblue"),
    risk.table = TRUE,  risk.table.height = 0.25, xscale = "d_y", break.time.by=365.25*2.5) + ggtitle("Cluster E")

survminer::ggsurvplot(
    fit = survival::survfit(BasesF ~ dana1, data = ForSurvivalF), 
    xlab = "Years", 
    ylab = "Overall survival probability",
    legend.title = "",
    palette=c("deeppink", "midnight blue", "cornflowerblue"),
    risk.table = TRUE,  risk.table.height = 0.25, xscale = "d_y", break.time.by=365.25*2.5) + ggtitle("Cluster F")

```

```{r}
survminer::ggsurvplot(
    fit = survival::survfit(Bases10 ~ dana1 + cluster6, data = ForSurvival), 
    xlab = "Years", 
    ylab = "Overall survival probability",
    legend.title = "",
    risk.table = TRUE,  risk.table.height = 0.25, xscale = "d_y", break.time.by=365.25*2.5)

```




# Clusters A, C & F

```{r}

ACFsurv <- ForSurvival %>% filter(cluster6 %in% c("A", "C", "F"))
ACFBases<- survival::Surv(time = ACFsurv$DaysToEvent10yr, event = ACFsurv$Survivor10yr)


survminer::ggsurvplot(
    fit = survival::survfit(ACFBases ~ cluster6, data = ACFsurv), 
    xlab = "Years", 
    ylab = "Overall survival probability",
    legend.title = "",
    palette=c("deeppink", "midnight blue", "cornflowerblue"), 
    risk.table = TRUE, , xscale = "d_y", break.time.by=365.25*2.5)

broom::tidy(coxph(ACFBases~cluster6, data = ACFsurv), exp = TRUE) %>% kable(., caption = "Cox regression: ~ clusters A, C & F") %>% kable_styling(full_width = TRUE)
coxph(ACFBases~cluster6, data = ACFsurv) %>% gtsummary::tbl_regression(exp = TRUE)

broom::tidy(coxph(ACFBases~sex+age+cluster6, data = ACFsurv), exp = TRUE) %>% kable(., caption = "Cox regression: ~ clusters A, B & E") %>% kable_styling(full_width = TRUE)
coxph(ACFBases~sex+age+cluster6, data = ACFsurv) %>% gtsummary::tbl_regression(exp = TRUE)

ggforest(coxph(ACFBases~sex+age+cluster6, data = ACFsurv))
```

## WHO functional class survival


```{r}


ForSurvival$fclass <- as.factor(ForSurvival$fclass)
survminer::ggsurvplot(
    fit = survival::survfit(Bases ~ fclass, data = ForSurvival), 
    xlab = "Years", 
    ylab = "Overall survival probability",
    legend.title = "",
    palette=c("deeppink","dark red", "midnight blue", "cornflowerblue"),
    risk.table = TRUE,  risk.table.height = 0.3, xscale = "d_y", break.time.by=365.25*2.5)


broom::tidy(coxph(Bases~sex+age+fclass, data = ForSurvival), exp = TRUE) %>% kable(., caption = "Cox regression: ~ sex + age + FC") %>% kable_styling(full_width = TRUE)
coxph(Bases~sex+age+fclass, data = ForSurvival) %>% gtsummary::tbl_regression(exp = TRUE)

broom::tidy(coxph(Bases~fclass, data = ForSurvival), exp = TRUE) %>% kable(., caption = "Cox regression: ~ FC") %>% kable_styling(full_width = TRUE)
coxph(Bases~fclass, data = ForSurvival) %>% gtsummary::tbl_regression(exp = TRUE)

ggforest(coxph(Bases~sex+age+fclass, data = ForSurvival))
```

## REVEAL group survival


```{r}
survminer::ggsurvplot(
    fit = survival::survfit(Bases ~ REVEALgroup, data = ForSurvival), 
    xlab = "Years", 
    ylab = "Overall survival probability",
    legend.title = "",
    palette=c("deeppink", "midnight blue", "cornflowerblue"),
    risk.table = TRUE,  risk.table.height = 0.3, xscale = "d_y", break.time.by=365.25*2.5)


broom::tidy(coxph(Bases~sex+age+REVEALgroup, data = ForSurvival), exp = TRUE) %>% kable(., caption = "Cox regression: ~ sex + age + REVEAL group") %>% kable_styling(full_width = TRUE)
coxph(Bases~sex+age+REVEALgroup, data = ForSurvival) %>% gtsummary::tbl_regression(exp = TRUE)

broom::tidy(coxph(Bases~REVEALgroup, data = ForSurvival), exp = TRUE) %>% kable(., caption = "Cox regression: ~ REVEAL group") %>% kable_styling(full_width = TRUE)
coxph(Bases~REVEALgroup, data = ForSurvival) %>% gtsummary::tbl_regression(exp = TRUE)

ggforest(coxph(Bases~sex+age+REVEALgroup, data = ForSurvival))
```



```{r}
ForSurvivalHigh <- ForSurvival %>% filter(REVEALgroup == "High")
BasesHigh<- survival::Surv(time = ForSurvivalHigh$DaysToEvent10yr, event = ForSurvivalHigh$Survivor10yr)
ForSurvivalInt <- ForSurvival %>% filter(REVEALgroup == "Intermediate")
BasesInt<- survival::Surv(time = ForSurvivalInt$DaysToEvent10yr, event = ForSurvivalInt$Survivor10yr)
ForSurvivalLow <- ForSurvival %>% filter(REVEALgroup == "Low")
BasesLow<- survival::Surv(time = ForSurvivalLow$DaysToEvent10yr, event = ForSurvivalLow$Survivor10yr)

survminer::ggsurvplot(
    fit = survival::survfit(BasesHigh ~ cluster6, data = ForSurvivalHigh), 
    xlab = "Years", 
    ylab = "Overall survival probability",
    legend.title = "",
    palette=c("deeppink","dark red", "darksalmon", "yellow", "midnight blue", "cornflowerblue"),
     xscale = "d_y", break.time.by=365.25*2.5) + ggtitle("High REVEAL group")

survminer::ggsurvplot(
    fit = survival::survfit(BasesInt ~ cluster6, data = ForSurvivalInt), 
    xlab = "Years", 
    ylab = "Overall survival probability",
    legend.title = "",
    palette=c("deeppink","dark red", "darksalmon", "yellow", "midnight blue", "cornflowerblue"),
     xscale = "d_y", break.time.by=365.25*2.5) + ggtitle("Intermediate REVEAL group")

survminer::ggsurvplot(
    fit = survival::survfit(BasesLow ~ cluster6, data = ForSurvivalLow), 
    xlab = "Years", 
    ylab = "Overall survival probability",
    legend.title = "",
    palette=c("deeppink","dark red", "darksalmon", "yellow", "midnight blue", "cornflowerblue"),
     xscale = "d_y", break.time.by=365.25*2.5) + ggtitle("Low REVEAL group")
```

```{r}
survminer::ggsurvplot(
    fit = survival::survfit(BasesA ~ REVEALgroup, data = ForSurvivalA), 
    xlab = "Years", 
    ylab = "Overall survival probability",
    legend.title = "",
    palette=c("deeppink", "midnight blue", "cornflowerblue"),
    risk.table = TRUE,  risk.table.height = 0.25, xscale = "d_y", break.time.by=365.25*2.5) + ggtitle("Cluster A REVEAL groups")

survminer::ggsurvplot(
    fit = survival::survfit(BasesC ~ REVEALgroup, data = ForSurvivalC), 
    xlab = "Years", 
    ylab = "Overall survival probability",
    legend.title = "",
    palette=c("deeppink", "midnight blue", "cornflowerblue"),
    risk.table = TRUE,  risk.table.height = 0.25, xscale = "d_y", break.time.by=365.25*2.5) + ggtitle("Cluster C REVEAL groups")

survminer::ggsurvplot(
    fit = survival::survfit(BasesF ~ REVEALgroup, data = ForSurvivalF), 
    xlab = "Years", 
    ylab = "Overall survival probability",
    legend.title = "",
    palette=c("deeppink", "midnight blue", "cornflowerblue"),
    risk.table = TRUE,  risk.table.height = 0.25, xscale = "d_y", break.time.by=365.25*2.5) + ggtitle("Cluster F REVEAL groups")
```


## Treatment

```{r}
ForSurvivalTreatT <- ForSurvival %>% filter(treated == "TRUE")
BasesTreatT<- survival::Surv(time = ForSurvivalTreatT$DaysToEvent10yr, event = ForSurvivalTreatT$Survivor10yr)
ForSurvivalTreatF <- ForSurvival %>% filter(treated == "FALSE")
BasesTreatF<- survival::Surv(time = ForSurvivalTreatF$DaysToEvent10yr, event = ForSurvivalTreatF$Survivor10yr)

survminer::ggsurvplot(
    fit = survival::survfit(BasesTreatT ~ cluster6, data = ForSurvivalTreatT), 
    xlab = "Years", 
    ylab = "Overall survival probability",
    legend.title = "",
    palette=c("deeppink","dark red", "darksalmon", "yellow", "midnight blue", "cornflowerblue"),
    risk.table = TRUE,  risk.table.height = 0.25, xscale = "d_y", break.time.by=365.25*2.5) + ggtitle("Treated patients by cluster")

survminer::ggsurvplot(
    fit = survival::survfit(BasesTreatF ~ cluster6, data = ForSurvivalTreatF), 
    xlab = "Years", 
    ylab = "Overall survival probability",
    legend.title = "",
    palette=c("deeppink","dark red", "darksalmon", "yellow", "midnight blue", "cornflowerblue"),
    risk.table = TRUE,  risk.table.height = 0.25, xscale = "d_y", break.time.by=365.25*2.5) + ggtitle("Treatment naive patients by cluster")
```

```{r}
survminer::ggsurvplot(
    fit = survival::survfit(BasesC ~ treated, data = ForSurvivalC), 
    xlab = "Years", 
    ylab = "Overall survival probability",
    legend.title = "",
    palette=c("midnight blue", "cornflowerblue"),
    risk.table = TRUE,  risk.table.height = 0.25, xscale = "d_y", break.time.by=365.25*2.5) + ggtitle("Cluster C by treatment status")
```

```{r}
survminer::ggsurvplot(
    fit = survival::survfit(BasesD ~ treated, data = ForSurvivalD), 
    xlab = "Years", 
    ylab = "Overall survival probability",
    legend.title = "",
    palette=c("midnight blue", "cornflowerblue"),
    risk.table = TRUE,  risk.table.height = 0.25, xscale = "d_y", break.time.by=365.25*2.5) + ggtitle("Cluster D by treatment status")
```


```{r}
ForSurvivalnoPH23 <- ForSurvival %>% filter(dana1 == "PH1")
BasesnoPH23<- survival::Surv(time = ForSurvivalnoPH23$DaysToEvent10yr, event = ForSurvivalnoPH23$Survivor10yr)

survminer::ggsurvplot(
    fit = survival::survfit(BasesnoPH23 ~ cluster6, data = ForSurvivalnoPH23), 
    xlab = "Years", 
    ylab = "Overall survival probability",
    legend.title = "",
    palette=c("deeppink","dark red", "darksalmon", "yellow", "midnight blue", "cornflowerblue"),
    risk.table = TRUE,  risk.table.height = 0.25, xscale = "d_y", break.time.by=365.25*2.5) + ggtitle("Cluster survival for PH1 only")
```

```{r}
ForSurvivalPH23 <- ForSurvival %>% filter(dana1 != "PH1")
BasesPH23<- survival::Surv(time = ForSurvivalPH23$DaysToEvent10yr, event = ForSurvivalPH23$Survivor10yr)

survminer::ggsurvplot(
    fit = survival::survfit(BasesPH23 ~ cluster6, data = ForSurvivalPH23), 
    xlab = "Years", 
    ylab = "Overall survival probability",
    legend.title = "",
    palette=c("deeppink","dark red", "darksalmon", "yellow", "midnight blue", "cornflowerblue"),
    risk.table = TRUE,  risk.table.height = 0.25, xscale = "d_y", break.time.by=365.25*2.5) + ggtitle("Cluster survival for PH2 & PH3")
```

## PH group

```{r}
survminer::ggsurvplot(
    fit = survival::survfit(Bases ~ dana, data = ForSurvival), 
    xlab = "Years", 
    ylab = "Overall survival probability",
    legend.title = "",
    palette=c("deeppink","dark red", "darksalmon", "yellow", "midnight blue", "cornflowerblue"),
    risk.table = TRUE,  risk.table.height = 0.25, xscale = "d_y", break.time.by=365.25*2.5)


broom::tidy(coxph(Bases~sex+age+dana, data = ForSurvival), exp = TRUE) %>% kable(., caption = "Cox regression: ~ sex + age + 6 clusters") %>% kable_styling(full_width = TRUE)
coxph(Bases~sex+age+dana, data = ForSurvival) %>% gtsummary::tbl_regression(exp = TRUE)

broom::tidy(coxph(Bases~dana, data = ForSurvival), exp = TRUE) %>% kable(., caption = "Cox regression: ~ 6 clusters") %>% kable_styling(full_width = TRUE)
coxph(Bases~dana, data = ForSurvival) %>% gtsummary::tbl_regression(exp = TRUE)

ggforest(coxph(Bases~sex+age+dana, data = ForSurvival))
```

# Missing data


```{r}
kable(freq.na(Clusterpheno), caption = "Missing values for all columns") %>% kable_styling(full_width = TRUE)
```

```{r}
kable(cbind(addmargins(table(rowSums(is.na(Clusterpheno))),1), addmargins(round(prop.table(table(rowSums(is.na(Clusterpheno))))*100, digits = 1),1)), caption = "Number of missing variables for each patient", col.names = c("n", "%")) %>% kable_styling(full_width = TRUE)
```


```{r, eval = F, include = F}
#save.image("Preprocessing.RData")
load("Preprocessing.RData")
#forcyto <- Clusterpheno %>% select(-cluster2, -cluster3, -cluster4, -cluster5, -cluster7, -cluster8, -cluster9, -cluster10)
#write_csv(forcyto, "PH1PH2PH3network.csv")
Clusterpheno %>% count(cluster6, fclass, REVEALgroup) %>% rownames_to_column() %>% write_csv("PH1PH2PH3network.csv")
```




# Correlations

```{r, warning=FALSE}
forcorr <- select(Clusterpheno, -cluster6, -PARTITION, -dana, -dana1, -dana1.1, -dana1.1.1, -dana1.1.1.1, -REVEALgroup, -thyroid_disease, -t2dm, -copd, -SleepApnoea, -ihd, -AtrialFibrillation, -runid, -OtherCTD, -ssc, -ERA, -Prostanoid, -PDE5, -Other, -sex, -site, - treated, -Ethnicity)

library(lares)

corr_cross(forcorr,  method = "spearman", # name of dataset
           contains = c("age", "bmi", "bpsys", "bpdia", "fclass", "smwd", "fvcp", "ntprobnp", "ua", "platelet_count", "eGFR", "mPAP", "mRAP", "PAWP", "dPAP", "sPAP", "SvO2", "tlco_predicted", "PVRdynes", "Cardiac Index", "Cardiac Output", "mRAP"),
  top = 100 # display top 10 couples of variables (by correlation coefficient)
)
```


```{r, warning=FALSE}
allcors<- corr_cross(forcorr, method = "spearman",  # name of dataset
           contains = c("age", "bmi", "bpsys", "bpdia", "fclass", "smwd", "fvcp", "ntprobnp", "ua", "platelet_count", "eGFR", "mPAP", "mRAP", "PAWP", "dPAP", "sPAP", "SvO2", "tlco_predicted", "PVRdynes", "Cardiac Index", "Cardiac Output", "mRAP"), type = 2
)

allcorsdata<- allcors$data 
allcorsdata %>% mutate(adjpval = round(p.adjust(pvalue, "BH"), 4)) %>% filter(adjpval < 0.05) %>% filter(size > 0.2) %>% filter(grepl("hsa", mix)) %>% select(key, mix, size, corr, pvalue, adjpval) %>% kable(., caption = "Correlation coefficients > 0.2 for clinical parameters & miRNAs, adjuster pvalue < 0.05") %>% kable_styling(full_width = TRUE)

```


## PVR

```{r, warning=FALSE}
corr_cross(forcorr,  method = "spearman", 
           type = 2,
           contains = "PVRdynes", max_pvalue = 0.05 )

corrPVR<- corr_cross(forcorr, # name of dataset,
           type = 1,
           contains = "PVRdynes", top = 100)

corrPVR<- corrPVR$data
corrPVR %>% filter(grepl("hsa", mix)) %>% select(Comparison = label, `correlation coeff` = corr, `absolute correlation coeff` = abs, pvalue) %>% mutate(`Adjusted Pvalue` = round(p.adjust(pvalue, "BH"), 4)) %>% filter(`Adjusted Pvalue` < 0.05) %>% kable(., caption = "miRNAs correlating with PVR, adjusted pvalue < 0.05") %>% kable_styling(full_width = TRUE)
```

## mPAP

```{r, warning=FALSE}
corr_cross(forcorr,  method = "spearman", 
           type = 2,
           contains = "mPAP", max_pvalue = 0.05 )

corrmPAP<- corr_cross(forcorr, # name of dataset,
           type = 1,
           contains = "mPAP", top = 100)

corrmPAP<- corrmPAP$data
corrmPAP %>%filter(grepl("hsa", mix)) %>% select(Comparison = label, `correlation coeff` = corr, `absolute correlation coeff` = abs, pvalue) %>% mutate(`Adjusted Pvalue` = round(p.adjust(pvalue, "BH"), 4)) %>% filter(`Adjusted Pvalue` < 0.05) %>% kable(., caption = "miRNAs correlating with mPAP, adjusted pvalue < 0.05") %>% kable_styling(full_width = TRUE)
```

## dPAP

```{r, warning=FALSE}
corr_cross(forcorr,  method = "spearman", 
           type = 2,
           contains = "dPAP", max_pvalue = 0.05 )

corrdPAP<- corr_cross(forcorr, # name of dataset,
           type = 1,
           contains = "dPAP", top = 100)

corrdPAP<- corrdPAP$data
corrdPAP %>% filter(grepl("hsa", mix)) %>% select(Comparison = label, `correlation coeff` = corr, `absolute correlation coeff` = abs, pvalue) %>% mutate(`Adjusted Pvalue` = round(p.adjust(pvalue, "BH"), 4))%>% filter(`Adjusted Pvalue` < 0.05) %>% kable(., caption = "miRNAs correlating with dPAP, adjusted pvalue < 0.05") %>% kable_styling(full_width = TRUE)
```

## mRAP

```{r, warning=FALSE}
corr_cross(forcorr,  method = "spearman", 
           type = 2,
           contains = "mRAP", max_pvalue = 0.05 )
corrmRAP<- corr_cross(forcorr, # name of dataset,
           type = 1,
           contains = "mRAP", top = 100)

corrmRAP<- corrmRAP$data
corrmRAP %>% filter(grepl("hsa", mix)) %>% select(Comparison = label, `correlation coeff` = corr, `absolute correlation coeff` = abs, pvalue) %>% mutate(`Adjusted Pvalue` = round(p.adjust(pvalue, "BH"), 4)) %>% filter(`Adjusted Pvalue` < 0.05) %>% kable(., caption = "miRNAs correlating with mRAP, adjusted pvalue < 0.05") %>% kable_styling(full_width = TRUE)
```

## PAWP

```{r, warning=FALSE}
corr_cross(forcorr,  method = "spearman", 
           type = 2,
           contains = "PAWP", max_pvalue = 0.05 )

corrPAWP<- corr_cross(forcorr, # name of dataset,
           type = 1,
           contains = "PAWP", top = 100)

corrPAWP<- corrPAWP$data
corrPAWP %>% filter(grepl("hsa", mix)) %>% select(Comparison = label, `correlation coeff` = corr, `absolute correlation coeff` = abs, pvalue) %>% mutate(`Adjusted Pvalue` = round(p.adjust(pvalue, "BH"), 4)) %>% filter(`Adjusted Pvalue` < 0.05) %>% kable(., caption = "miRNAs correlating with PAWP, adjusted pvalue < 0.05") %>% kable_styling(full_width = TRUE)
```

## Cardiac Index

```{r, warning=FALSE}
corr_cross(forcorr, method = "spearman",  
           type = 2,
           contains = "CardiacIndex", max_pvalue = 0.05 )

corrCardiacIndex<- corr_cross(forcorr, # name of dataset,
           type = 1,
           contains = "CardiacIndex", top = 100)

corrCardiacIndex<- corrCardiacIndex$data
corrCardiacIndex %>% filter(grepl("hsa", mix)) %>% select(Comparison = label, `correlation coeff` = corr, `absolute correlation coeff` = abs, pvalue) %>% mutate(`Adjusted Pvalue` = round(p.adjust(pvalue, "BH"), 4)) %>% filter(`Adjusted Pvalue` < 0.05) %>% kable(., caption = "miRNAs correlating with Cardiac Index, adjusted pvalue < 0.05") %>% kable_styling(full_width = TRUE)
```

## Cardiac Output

```{r, warning=FALSE}
corr_cross(forcorr,  method = "spearman", 
           type = 2,
           contains = "CardiacOutput", max_pvalue = 0.05 )

corrCardiacOutput<- corr_cross(forcorr, # name of dataset,
           type = 1,
           contains = "CardiacOutput", top = 100)

corrCardiacOutput<- corrCardiacOutput$data
corrCardiacOutput %>% filter(grepl("hsa", mix)) %>% select(Comparison = label, `correlation coeff` = corr, `absolute correlation coeff` = abs, pvalue) %>% mutate(`Adjusted Pvalue` = round(p.adjust(pvalue, "BH"), 4)) %>% filter(`Adjusted Pvalue` < 0.05) %>% kable(., caption = "miRNAs correlating with Cardiac Output, adjusted pvalue < 0.05") %>% kable_styling(full_width = TRUE)
```

## SvO2

```{r, warning=FALSE}
corr_cross(forcorr,  method = "spearman", 
           type = 2,
           contains = "SvO2", max_pvalue = 0.05 )

corrSvO2<- corr_cross(forcorr, # name of dataset,
           type = 1,
           contains = "SvO2", top = 100)

corrSvO2<- corrSvO2$data
corrSvO2 %>% filter(grepl("hsa", mix)) %>% select(Comparison = label, `correlation coeff` = corr, `absolute correlation coeff` = abs, pvalue) %>% mutate(`Adjusted Pvalue` = round(p.adjust(pvalue, "BH"), 4)) %>% filter(`Adjusted Pvalue` < 0.05) %>% kable(., caption = "miRNAs correlating with SvO2, adjusted pvalue < 0.05") %>% kable_styling(full_width = TRUE)
```

## eGFR

```{r, warning=FALSE}
corr_cross(forcorr,  method = "spearman", 
           type = 2,
           contains = "eGFR", max_pvalue = 0.05 )

coreGFR<- corr_cross(forcorr, # name of dataset,
           type = 1,
           contains = "eGFR", top = 100)

coreGFR<- coreGFR$data
coreGFR %>% filter(grepl("hsa", mix)) %>% select(Comparison = label, `correlation coeff` = corr, `absolute correlation coeff` = abs, pvalue) %>% mutate(`Adjusted Pvalue` = round(p.adjust(pvalue, "BH"), 4)) %>% filter(`Adjusted Pvalue` < 0.05) %>% kable(., caption = "miRNAs correlating with eGFR, adjusted pvalue < 0.05") %>% kable_styling(full_width = TRUE)
```

## TLco predicted

```{r, warning=FALSE}
corr_cross(forcorr,  method = "spearman", 
           type = 2,
           contains = "tlco_predicted", max_pvalue = 0.05 )

corrtlco <- corr_cross(forcorr, # name of dataset,
           type = 1,
           contains = "tlco_predicted", top = 100)

corrtlco<- corrtlco$data
corrtlco %>% filter(grepl("hsa", mix)) %>% select(Comparison = label, `correlation coeff` = corr, `absolute correlation coeff` = abs, pvalue) %>% mutate(`Adjusted Pvalue` = round(p.adjust(pvalue, "BH"), 4)) %>% filter(`Adjusted Pvalue` < 0.05) %>% kable(., caption = "miRNAs correlating with tlco predicted, adjusted pvalue < 0.05") %>% kable_styling(full_width = TRUE)
```

## BP-systolic

```{r, warning=FALSE}
corr_cross(forcorr,  method = "spearman", 
           type = 2,
           contains = "bpsys", max_pvalue = 0.05 )

corrbpsys<- corr_cross(forcorr, # name of dataset,
           type = 1,
           contains = "bpsys", top = 100)

corrbpsys<- corrbpsys$data
corrbpsys %>% filter(grepl("hsa", mix)) %>% select(Comparison = label, `correlation coeff` = corr, `absolute correlation coeff` = abs, pvalue) %>% mutate(`Adjusted Pvalue` = round(p.adjust(pvalue, "BH"), 4)) %>% filter(`Adjusted Pvalue` < 0.05) %>% kable(., caption = "miRNAs correlating with bpsys, adjusted pvalue < 0.05") %>% kable_styling(full_width = TRUE)
```

## BP-diastolic

```{r, warning=FALSE}
corr_cross(forcorr, method = "spearman", 
           type = 2,
           contains = "bpdia", max_pvalue = 0.05 )

corrbpdia<- corr_cross(forcorr, # name of dataset,
           type = 1,
           contains = "bpdia", top = 100)

corrbpdia<- corrbpdia$data
corrbpdia %>% filter(grepl("hsa", mix)) %>% select(Comparison = label, `correlation coeff` = corr, `absolute correlation coeff` = abs, pvalue) %>% mutate(`Adjusted Pvalue` = round(p.adjust(pvalue, "BH"), 4)) %>% filter(`Adjusted Pvalue` < 0.05) %>% kable(., caption = "miRNAs correlating with bpdia, adjusted pvalue < 0.05") %>% kable_styling(full_width = TRUE)
```

## Functional class

```{r, warning=FALSE}
corr_cross(forcorr,  method = "spearman", 
           type = 2,
           contains = "fclass", max_pvalue = 0.05 )

corrfclass<- corr_cross(forcorr, # name of dataset,
           type = 1,
           contains = "fclass", top = 100)

corrfclass<- corrfclass$data
corrfclass %>% filter(grepl("hsa", mix)) %>% select(Comparison = label, `correlation coeff` = corr, `absolute correlation coeff` = abs, pvalue) %>% mutate(`Adjusted Pvalue` = round(p.adjust(pvalue, "BH"), 4)) %>% filter(`Adjusted Pvalue` < 0.05)  %>% kable(., caption = "miRNAs correlating with fclass, adjusted pvalue < 0.05") %>% kable_styling(full_width = TRUE)
```

## 6MWD

```{r, warning=FALSE}
corr_cross(forcorr,  method = "spearman", 
           type = 2,
           contains = "smwd", max_pvalue = 0.05 )

corrsmwd<- corr_cross(forcorr, # name of dataset,
           type = 1,
           contains = "smwd", top = 100)

corrsmwd<- corrsmwd$data
corrsmwd %>% filter(grepl("hsa", mix)) %>% select(Comparison = label, `correlation coeff` = corr, `absolute correlation coeff` = abs, pvalue) %>% mutate(`Adjusted Pvalue` = round(p.adjust(pvalue, "BH"), 4)) %>% filter(`Adjusted Pvalue` < 0.05) %>% kable(., caption = "miRNAs correlating with 6mwd, adjusted pvalue < 0.05") %>% kable_styling(full_width = TRUE)
```

## NT-proBNP

```{r, warning=FALSE}
corr_cross(forcorr,  method = "spearman", 
           type = 2,
           contains = "ntprobnp", max_pvalue = 0.05 )

corrntprobnp<- corr_cross(forcorr, # name of dataset,
           type = 1,
           contains = "ntprobnp", top = 100)

corrntprobnp<- corrntprobnp$data
corrntprobnp %>% filter(grepl("hsa", mix)) %>% select(Comparison = label, `correlation coeff` = corr, `absolute correlation coeff` = abs, pvalue) %>% mutate(`Adjusted Pvalue` = round(p.adjust(pvalue, "BH"), 4)) %>% filter(`Adjusted Pvalue` < 0.05) %>% kable(., caption = "miRNAs correlating with ntprobnp, adjusted pvalue < 0.05") %>% kable_styling(full_width = TRUE)
```

## Uric Acid

```{r, warning=FALSE}
corr_cross(forcorr,  method = "spearman", 
           type = 2,
           contains = "ua", max_pvalue = 0.05 )

corrua<- corr_cross(forcorr, # name of dataset,
           type = 1,
           contains = "ua", top = 100)

corrua<- corrua$data
corrua %>% filter(grepl("hsa", label)) %>% select(Comparison = label, `correlation coeff` = corr, `absolute correlation coeff` = abs, pvalue) %>% mutate(`Adjusted Pvalue` = round(p.adjust(pvalue, "BH"), 4)) %>% filter(`Adjusted Pvalue` < 0.05) %>% kable(., caption = "miRNAs correlating with ua, adjusted pvalue < 0.05") %>% kable_styling(full_width = TRUE)
```

# WHO classification groups

```{r}
Clusterpheno %>% group_by(dana) %>% summarise(mean = mean(ntprobnp), sd = sd(ntprobnp)) %>% kable(., caption = "NT-proBNP by WHO classificaiton, all clusters") %>% kable_styling(full_width = TRUE)
```

# ACF comparisons

```{r}
as.data.frame(table(Clusterpheno$dana, Clusterpheno$cluster6)) %>% filter(Var2 %in% c("A", "C","F")) %>% group_by(Var2, Var1) %>% ggplot(aes(x="", y=Freq, fill=Var1)) + geom_bar(stat="identity", width = 1, color = "white") + cp + theme_void()  + facet_wrap(~Var2, scales = "free") + theme(aspect.ratio = 1) + scale_fill_manual(values = c("deeppink","dark red", "darksalmon", "yellow", "midnight blue", "cornflowerblue"))

as.data.frame(table(Clusterpheno$dana, Clusterpheno$cluster6)) %>% filter(Var2 %in% c("A", "C","F"))  %>% group_by(Var2, Var1) %>% ggplot(aes(x=Var2, y=Freq, fill=Var1)) + geom_bar(stat="identity")  + scale_fill_manual(values = c("deeppink","dark red", "darksalmon", "yellow", "midnight blue", "cornflowerblue")) + ggtitle("PH Status, 6 Clusters") + labs(x = "Cluster", y = "") +  guides(fill=guide_legend(title=NULL))

```


```{r}
as.data.frame(table(Clusterpheno$fclass, Clusterpheno$cluster6)) %>% filter(Var2 %in% c("A", "C","F")) %>% group_by(Var2, Var1) %>% ggplot(aes(x="", y=Freq, fill=Var1)) + geom_bar(stat="identity", width = 1, color = "white") + cp + theme_void()  + facet_wrap(~Var2, scales = "free") + theme(aspect.ratio = 1) + scale_fill_manual(values = c("deeppink", "darksalmon", "midnight blue", "cornflowerblue"))  +  guides(fill=guide_legend(title='Functional Class'))

as.data.frame(table(Clusterpheno$fclass, Clusterpheno$cluster6)) %>% filter(Var2 %in% c("A", "C","F"))  %>% group_by(Var2, Var1) %>% ggplot(aes(x=Var2, y=Freq, fill=Var1)) + geom_bar(stat="identity")  + scale_fill_manual(values = c("deeppink","darksalmon","midnight blue", "cornflowerblue")) + ggtitle("Functional class, Clusters A, C & F") + labs(x = "Cluster", y = "n") +  guides(fill=guide_legend(title='Functional Class'))

```

```{r}
as.data.frame(table(Clusterpheno$REVEALgroup, Clusterpheno$cluster6)) %>% filter(Var2 %in% c("A", "C","F")) %>% group_by(Var2, Var1) %>% ggplot(aes(x="", y=Freq, fill=Var1)) + geom_bar(stat="identity", width = 1, color = "white") + cp + theme_void()  + facet_wrap(~Var2, scales = "free") + theme(aspect.ratio = 1) + scale_fill_manual(values = c("deeppink","midnight blue", "cornflowerblue")) +  guides(fill=guide_legend(title='REVEAL risk group')) 

as.data.frame(table(Clusterpheno$REVEALgroup, Clusterpheno$cluster6)) %>% filter(Var2 %in% c("A", "C","F"))  %>% group_by(Var2, Var1) %>% ggplot(aes(x=Var2, y=Freq, fill=Var1)) + geom_bar(stat="identity")  + scale_fill_manual(values = c("deeppink","midnight blue", "cornflowerblue")) + ggtitle("REVEAL group, Clusters A, C & F") + labs(x = "Cluster", y = "") +  guides(fill=guide_legend(title='REVEAL risk group')) 

```

```{r}
survmeasures<- Clusterpheno %>% filter(cluster6 %in% c("A", "C", "F")) %>% select(cluster6, fclass, REVEALgroup)
FCprop <- as.data.frame(prop.table(table(survmeasures$cluster6, survmeasures$fclass),1)*100)
revealprop <- as.data.frame(prop.table(table(survmeasures$cluster6, survmeasures$REVEALgroup),1)*100)
survmeasureprop <- rbind(FCprop, revealprop)
survmeasureprop$vartype <- ifelse(survmeasureprop$Var2 %in% c("High", "Intermediate", "Low"), "Functional class", "REVEAL risk group")

survmeasureprop %>% ggplot(aes(x = Var1, y = Freq, fill = Var2)) + geom_bar(stat = 'identity', position = 'dodge') + facet_wrap(~vartype) + scale_fill_manual(values = c("midnight blue","cornflowerblue","darksalmon", "deeppink", "deeppink","darksalmon", "midnight blue")) + labs(x = "Cluster", y = "% composition ") + theme(text = element_text(size = 14), panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey")) +  guides(fill=guide_legend(title=NULL))

survmeasures %>% count(cluster6, fclass, REVEALgroup) %>% melt(., id.vars = c('cluster6', 'n')) %>% filter(!is.na(value)) %>% ggplot(aes(x = cluster6, y = n, fill = value)) + geom_bar(stat = 'identity', position = 'dodge') + facet_wrap(~variable, labeller = labeller(variable = c("fclass" = "Functional class", "REVEALgroup" = "REVEAL risk group"))) + scale_fill_manual(values = c("midnight blue","cornflowerblue","darksalmon", "deeppink", "deeppink","darksalmon", "midnight blue")) + labs(x = "Cluster", y = "n") + theme(text = element_text(size = 14), panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey")) +  guides(fill=guide_legend(title=NULL))

survmeasures %>% count(cluster6, fclass, REVEALgroup) %>% melt(., id.vars = c('cluster6', 'n')) %>% filter(value != NA) %>% ggplot(aes(x = cluster6, y = n, fill = cluster6, pattern = value))+ scale_fill_manual(values = c("deeppink", "darksalmon",  "cornflowerblue"))  + geom_bar_pattern(stat = "identity", position = "dodge", color = "black", 
  pattern_fill = "black",
                   pattern_angle = 45,
                   pattern_density = 0.1,
                   pattern_spacing = 0.025)+  scale_pattern_manual(values = c(Low = "stripe", Intermediate = "circle", High = "none", `1` = "none", `2` = "crosshatch", `3`= "stripe",`4` = "circle")) +  ggtitle("") + labs(x = "Cluster", y = "") + guides(fill=guide_legend(title=NULL)) + theme(panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey")) 
```

# No outlier figures for paper

```{r}
#function to re-adjust scale for boxplots when outliers are removed
calc_boxplot_stat <- function(x) {
  coef <- 1.5
  n <- sum(!is.na(x))
  # calculate quantiles
  stats <- quantile(x, probs = c(0.0, 0.25, 0.5, 0.75, 1.0))
  names(stats) <- c("ymin", "lower", "middle", "upper", "ymax")
  iqr <- diff(stats[c(2, 4)])
  # set whiskers

  outliers <- x < (stats[2] - coef * iqr) | x > (stats[4] + coef * iqr)
  if (any(outliers)) {
    stats[c(1, 5)] <- range(c(stats[2:4], x[!outliers]), na.rm = TRUE)
  }
  return(stats)
}

#Clusterpheno  %>%  select(cluster6, sPAP, mRAP) %>% melt(., id.vars = c('cluster6')) %>% filter(!is.na(value)) %>% ggplot(aes(x = cluster6, y = value, fill = cluster6)) + stat_summary(fun.data = calc_boxplot_stat, geom = "boxplot") + facet_wrap(~variable, scale = "free_y") + scale_fill_manual(values = clusterfills) + labs(x = "Cluster", y = "") + theme(text = element_text(size = 14), legend.position = "none", panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey")) +  guides(fill=guide_legend(title=NULL))

Clusterpheno %>% select(cluster6, mPAP) %>% filter(!is.na(mPAP)) %>% ggplot(aes(x= cluster6, y =mPAP, fill = cluster6)) + labs(x = "Cluster", y = "mPAP (mm Hg)") + theme(text = element_text(size = 14), legend.position = "none", panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey")) + stat_summary(fun.data = calc_boxplot_stat, geom = "boxplot")  + scale_fill_manual(values = clusterfills) +  guides(fill=guide_legend(title=NULL))

Clusterpheno %>% select(cluster6, PAWP) %>% filter(!is.na(PAWP)) %>% ggplot(aes(x= cluster6, y =PAWP, fill = cluster6)) + labs(x = "Cluster", y = "PAWP") + theme(text = element_text(size = 14), legend.position = "none", panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey")) + stat_summary(fun.data = calc_boxplot_stat, geom = "boxplot")  + scale_fill_manual(values = clusterfills) +  guides(fill=guide_legend(title=NULL))

Clusterpheno %>% select(cluster6, CardiacIndex) %>% filter(!is.na(CardiacIndex)) %>% ggplot(aes(x= cluster6, y =CardiacIndex, fill = cluster6)) + labs(x = "Cluster", y = "Cardiac Index") + theme(text = element_text(size = 14), legend.position = "none", panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey")) + stat_summary(fun.data = calc_boxplot_stat, geom = "boxplot")  + scale_fill_manual(values = clusterfills) +  guides(fill=guide_legend(title=NULL))

Clusterpheno %>% select(cluster6, CardiacOutput) %>% filter(!is.na(CardiacOutput)) %>% ggplot(aes(x= cluster6, y =CardiacOutput, fill = cluster6)) + labs(x = "Cluster", y = "Cardiac Output") + theme(text = element_text(size = 14), legend.position = "none", panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey")) + stat_summary(fun.data = calc_boxplot_stat, geom = "boxplot")  + scale_fill_manual(values = clusterfills) +  guides(fill=guide_legend(title=NULL))

Clusterpheno %>% select(cluster6, PVRdynes) %>% filter(!is.na(PVRdynes)) %>% ggplot(aes(x= cluster6, y =PVRdynes, fill = cluster6)) + labs(x = "Cluster", y = "PVR (Dynes)") + theme(text = element_text(size = 14), legend.position = "none", panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey")) + stat_summary(fun.data = calc_boxplot_stat, geom = "boxplot")  + scale_fill_manual(values = clusterfills) +  guides(fill=guide_legend(title=NULL))

Clusterpheno %>% select(cluster6, smwd) %>% filter(!is.na(smwd)) %>% ggplot(aes(x= cluster6, y =smwd, fill = cluster6)) + labs(x = "Cluster", y = "6MWD") + theme(text = element_text(size = 14), legend.position = "none", panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey")) + stat_summary(fun.data = calc_boxplot_stat, geom = "boxplot")  + scale_fill_manual(values = clusterfills) +  guides(fill=guide_legend(title=NULL))

Clusterpheno %>% select(cluster6, tlco_predicted) %>% filter(!is.na(tlco_predicted)) %>% ggplot(aes(x= cluster6, y =tlco_predicted, fill = cluster6)) + labs(x = "Cluster", y = "TLCO predicted") + theme(text = element_text(size = 14), legend.position = "none", panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey")) + stat_summary(fun.data = calc_boxplot_stat, geom = "boxplot")  + scale_fill_manual(values = clusterfills) +  guides(fill=guide_legend(title=NULL))

Clusterpheno %>% select(cluster6, SvO2) %>% filter(!is.na(SvO2)) %>% ggplot(aes(x= cluster6, y =SvO2, fill = cluster6)) + labs(x = "Cluster", y = "SvO2") + theme(text = element_text(size = 14), legend.position = "none", panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey")) + stat_summary(fun.data = calc_boxplot_stat, geom = "boxplot")  + scale_fill_manual(values = clusterfills) +  guides(fill=guide_legend(title=NULL))

Clusterpheno$origntprobnp <- 2^Clusterpheno$ntprobnp

Clusterpheno %>% select(cluster6, origntprobnp) %>% filter(!is.na(origntprobnp )) %>% ggplot(aes(x= cluster6, y =origntprobnp, fill = cluster6)) + labs(x = "Cluster", y = "NT-proBNP") + theme(text = element_text(size = 14), legend.position = "none", panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey")) + stat_summary(fun.data = calc_boxplot_stat, geom = "boxplot")  + scale_fill_manual(values = clusterfills) +  guides(fill=guide_legend(title=NULL))

```

# Significant clinical parameters

```{r}
Clusterpheno %>% select(cluster6, treated, site) %>% count(cluster6, treated, site) %>% melt(., id.vars = c('cluster6', 'n')) %>% ggplot(aes(x = cluster6, y = n, fill = value)) + geom_bar(stat = 'identity', position = 'dodge') + facet_wrap(~variable, labeller = labeller(variable = c("treated" = "Treatment status", "site" = "Site"))) + scale_fill_manual(values = c("deeppink","cornflowerblue","midnight blue", "cornflowerblue", "deeppink")) + labs(x = "Cluster", y = "n") + theme(text = element_text(size = 14), panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey")) +  guides(fill=guide_legend(title=NULL)) 

Clusterpheno %>% select(cluster6, mPAP, dPAP) %>% melt(., id.vars = c('cluster6')) %>% filter(!is.na(value)) %>% ggplot(aes(x = cluster6, y = value, fill = cluster6)) + stat_summary(fun.data = calc_boxplot_stat, geom = "boxplot")  + facet_wrap(~variable, scale ="free_y") + scale_fill_manual(values = c("deeppink","dark red", "darksalmon", "yellow", "midnight blue", "cornflowerblue")) + labs(x = "Cluster", y = "") + theme(text = element_text(size = 14), legend.position = "none", panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey")) +  guides(fill=guide_legend(title=NULL))

Clusterpheno %>% select(cluster6, sPAP, mRAP) %>% melt(., id.vars = c('cluster6')) %>% filter(!is.na(value)) %>% ggplot(aes(x = cluster6, y = value, fill = cluster6)) + stat_summary(fun.data = calc_boxplot_stat, geom = "boxplot") + facet_wrap(~variable, scale = "free_y") + scale_fill_manual(values = c("deeppink","dark red", "darksalmon", "yellow", "midnight blue", "cornflowerblue")) + labs(x = "Cluster", y = "") + theme(text = element_text(size = 14), legend.position = "none", panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey")) +  guides(fill=guide_legend(title=NULL))

Clusterpheno %>% select(cluster6, ntprobnp, Creatinine) %>% melt(., id.vars = c('cluster6')) %>% filter(!is.na(value)) %>% ggplot(aes(x = cluster6, y = value, fill = cluster6)) + stat_summary(fun.data = calc_boxplot_stat, geom = "boxplot") + facet_wrap(~variable, scale = "free_y", labeller = labeller(variable = c("ntprobnp" = "NT-proBNP", "Creatinine" = "Creatinine"))) + scale_fill_manual(values = c("deeppink","dark red", "darksalmon", "yellow", "midnight blue", "cornflowerblue")) + labs(x = "Cluster", y = "") + theme(text = element_text(size = 14), legend.position = "none", panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey")) +  guides(fill=guide_legend(title=NULL))

Clusterpheno %>% select(cluster6, SvO2, PVRdynes) %>% melt(., id.vars = c('cluster6')) %>% filter(!is.na(value)) %>% ggplot(aes(x = cluster6, y = value, fill = cluster6))+ stat_summary(fun.data = calc_boxplot_stat, geom = "boxplot") + facet_wrap(~variable, scale = "free_y", labeller = labeller(variable = c("SvO2" = "SvO2", "PVRdynes" = "PVR"))) + scale_fill_manual(values = c("deeppink","dark red", "darksalmon", "yellow", "midnight blue", "cornflowerblue")) + labs(x = "Cluster", y = "") + theme(text = element_text(size = 14), legend.position = "none", panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey")) +  guides(fill=guide_legend(title=NULL))

Clusterpheno %>% select(cluster6,  mPAP, dPAP, sPAP, mRAP, ntprobnp, Creatinine, SvO2, PVRdynes) %>% melt(., id.vars = c('cluster6')) %>% filter(!is.na(value)) %>% ggplot(aes(x = cluster6, y = value, fill = cluster6))+ stat_summary(fun.data = calc_boxplot_stat, geom = "boxplot") + facet_wrap(~variable, scales = "free") + scale_fill_manual(values = c("deeppink","dark red", "darksalmon", "yellow", "midnight blue", "cornflowerblue")) + labs(x = "Cluster", y = "") + theme(text = element_text(size = 14), legend.position = "none", panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey")) +  guides(fill=guide_legend(title=NULL))

#+ facet_grid(~variable, scale = "free_y", labeller = labeller(variable = c("SvO2" = "SvO2", "PVRdynes" = "PVR"))) 
```

```{r}
Clusterpheno %>% select(cluster6,  mPAP) %>% melt(., id.vars = c('cluster6')) %>% filter(!is.na(value)) %>% ggplot(aes(x = cluster6, y = value, fill = cluster6))+ stat_summary(fun.data = calc_boxplot_stat, geom = "boxplot") + scale_fill_manual(values = c("deeppink","dark red", "darksalmon", "yellow", "midnight blue", "cornflowerblue")) + labs(x = "Cluster", y = "") + theme(text = element_text(size = 14), legend.position = "none", panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey")) +  guides(fill=guide_legend(title=NULL)) + labs(y = "mPAP (mm Hg)")

Clusterpheno %>% select(cluster6,  dPAP) %>% melt(., id.vars = c('cluster6')) %>% filter(!is.na(value)) %>% ggplot(aes(x = cluster6, y = value, fill = cluster6))+ stat_summary(fun.data = calc_boxplot_stat, geom = "boxplot") + scale_fill_manual(values = c("deeppink","dark red", "darksalmon", "yellow", "midnight blue", "cornflowerblue")) + labs(x = "Cluster", y = "") + theme(text = element_text(size = 14), legend.position = "none", panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey")) +  guides(fill=guide_legend(title=NULL)) + labs(y = "dPAP (mm Hg)")

Clusterpheno %>% select(cluster6,  sPAP) %>% melt(., id.vars = c('cluster6')) %>% filter(!is.na(value)) %>% ggplot(aes(x = cluster6, y = value, fill = cluster6))+ stat_summary(fun.data = calc_boxplot_stat, geom = "boxplot") + scale_fill_manual(values = c("deeppink","dark red", "darksalmon", "yellow", "midnight blue", "cornflowerblue")) + labs(x = "Cluster", y = "") + theme(text = element_text(size = 14), legend.position = "none", panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey")) +  guides(fill=guide_legend(title=NULL)) + labs(y = "sPAP (mm Hg)")

Clusterpheno %>% select(cluster6, mRAP) %>% melt(., id.vars = c('cluster6')) %>% filter(!is.na(value)) %>% ggplot(aes(x = cluster6, y = value, fill = cluster6))+ stat_summary(fun.data = calc_boxplot_stat, geom = "boxplot") + scale_fill_manual(values = c("deeppink","dark red", "darksalmon", "yellow", "midnight blue", "cornflowerblue")) + labs(x = "Cluster", y = "") + theme(text = element_text(size = 14), legend.position = "none", panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey")) +  guides(fill=guide_legend(title=NULL)) + labs(y = "mRAP (mm Hg)")

Clusterpheno %>% select(cluster6, ntprobnp) %>% melt(., id.vars = c('cluster6')) %>% filter(!is.na(value)) %>% ggplot(aes(x = cluster6, y = value, fill = cluster6))+ stat_summary(fun.data = calc_boxplot_stat, geom = "boxplot") + scale_fill_manual(values = c("deeppink","dark red", "darksalmon", "yellow", "midnight blue", "cornflowerblue")) + labs(x = "Cluster", y = "") + theme(text = element_text(size = 14), legend.position = "none", panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey")) +  guides(fill=guide_legend(title=NULL)) + labs(y = "NT-proBNP ()")

Clusterpheno %>% select(cluster6, Creatinine) %>% melt(., id.vars = c('cluster6')) %>% filter(!is.na(value)) %>% ggplot(aes(x = cluster6, y = value, fill = cluster6))+ stat_summary(fun.data = calc_boxplot_stat, geom = "boxplot") + scale_fill_manual(values = c("deeppink","dark red", "darksalmon", "yellow", "midnight blue", "cornflowerblue")) + labs(x = "Cluster", y = "") + theme(text = element_text(size = 14), legend.position = "none", panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey")) +  guides(fill=guide_legend(title=NULL)) + labs(y = "Creatinine ()")

Clusterpheno %>% select(cluster6, SvO2) %>% melt(., id.vars = c('cluster6')) %>% filter(!is.na(value)) %>% ggplot(aes(x = cluster6, y = value, fill = cluster6))+ stat_summary(fun.data = calc_boxplot_stat, geom = "boxplot") + scale_fill_manual(values = c("deeppink","dark red", "darksalmon", "yellow", "midnight blue", "cornflowerblue")) + labs(x = "Cluster", y = "") + theme(text = element_text(size = 14), legend.position = "none", panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey")) +  guides(fill=guide_legend(title=NULL)) + labs(y = "SvO2 (%)")

Clusterpheno %>% select(cluster6, PVRdynes) %>% melt(., id.vars = c('cluster6')) %>% filter(!is.na(value)) %>% ggplot(aes(x = cluster6, y = value, fill = cluster6))+ stat_summary(fun.data = calc_boxplot_stat, geom = "boxplot") + scale_fill_manual(values = c("deeppink","dark red", "darksalmon", "yellow", "midnight blue", "cornflowerblue")) + labs(x = "Cluster", y = "") + theme(text = element_text(size = 14), legend.position = "none", panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey")) +  guides(fill=guide_legend(title=NULL)) + labs(y = "PVR (Dynes)")
```


```{r}
Clusterpheno %>% select(cluster6, smwd) %>% ggplot(aes(x= cluster6, y =smwd)) + geom_boxplot(fill="cornflowerblue") + labs(x = "Cluster", y = "6MWD") + theme(text = element_text(size = 14), legend.position = "none", panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey")) +  guides(fill=guide_legend(title=NULL))

Clusterpheno %>% select(cluster6, fvcp) %>% melt(., id.vars = c('cluster6')) %>% filter(!is.na(value)) %>% ggplot(aes(x = cluster6, y = value, fill = cluster6))+ stat_summary(fun.data = calc_boxplot_stat, geom = "boxplot") + scale_fill_manual(values = c("deeppink","dark red", "darksalmon", "yellow", "midnight blue", "cornflowerblue")) + labs(x = "Cluster", y = "") + theme(text = element_text(size = 14), legend.position = "none", panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey")) +  guides(fill=guide_legend(title=NULL)) + labs(y = "FVC (% Pred)")

Clusterpheno %>% select(cluster6, PAWP) %>% melt(., id.vars = c('cluster6')) %>% filter(!is.na(value)) %>% ggplot(aes(x = cluster6, y = value, fill = cluster6))+ stat_summary(fun.data = calc_boxplot_stat, geom = "boxplot") + scale_fill_manual(values = c("deeppink","dark red", "darksalmon", "yellow", "midnight blue", "cornflowerblue")) + labs(x = "Cluster", y = "") + theme(text = element_text(size = 14), legend.position = "none", panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey")) +  guides(fill=guide_legend(title=NULL)) + labs(y = "PAWP")
```

```{r}
Clusterpheno %>% select(site, smwd) %>% ggplot(aes(x= site, y =smwd)) + geom_boxplot(fill="cornflowerblue") + labs(x = "Cluster", y = "6MWD") + theme(text = element_text(size = 14), legend.position = "none", panel.background = element_blank(), panel.grid.major = element_line(colour = "grey"), axis.line = element_line(colour = "grey")) +  guides(fill=guide_legend(title=NULL))
```

# BWH data

```{r}
peak <- read_excel("/Volumes/GoogleDrive/Shared drives/MIREXS & Metabolon Retrospective Biomarker Study/BWH miRNA data/peak.xlsx")
post <- read_excel("/Volumes/GoogleDrive/Shared drives/MIREXS & Metabolon Retrospective Biomarker Study/BWH miRNA data/post.xlsx")
rest <- read_excel("/Volumes/GoogleDrive/Shared drives/MIREXS & Metabolon Retrospective Biomarker Study/BWH miRNA data/rest.xlsx")

#swap - for . in miR names to match
colnames(peak) <- gsub("-", ".", colnames(peak))
colnames(post) <- gsub("-", ".", colnames(post))
colnames(rest) <- gsub("-", ".", colnames(rest))

table(peak$dana) %>% kable(., caption = "DANA breakdown for peak") %>% kable_styling(full_width = TRUE)
table(post$dana) %>% kable(., caption = "DANA breakdown for post") %>% kable_styling(full_width = TRUE)
table(rest$dana) %>% kable(., caption = "DANA breakdown for rest") %>% kable_styling(full_width = TRUE)

```

# Diagnostically challenging patients

Diagnostically challenging patients: yes - patient incorrectly classified in >50% of CV folds

Treatment status -  False: patients PH treatment naive, True: patients on PH treatment at sampling

```{r, message = FALSE}
PH3vsPH1PH2classifications <- read_csv("PH3vsPH1PH2classifications.csv") %>% rename(PH3FPFNbinary = FPFNbinary)
PH2vsPH1PH3classifications <- read_csv("PH2vsPH1PH3classifications.csv") %>% rename(PH2FPFNbinary = FPFNbinary)
PH1vsPH2PH3classifications <- read_csv("PH1vsPH2PH3classifications.csv") %>% rename(PH1FPFNbinary = FPFNbinary)

ClusterphenoFPFN <- left_join(Clusterpheno, PH1vsPH2PH3classifications, by = "SampleID") %>% left_join(., PH2vsPH1PH3classifications, by = "SampleID") %>% left_join(., PH3vsPH1PH2classifications, by = "SampleID")

ClusterphenoFPFN %>% group_by(treated) %>% count(PH1FPFNbinary, PH2FPFNbinary, PH3FPFNbinary) %>% kable(.) %>% kable_styling(full_width = TRUE)

missclassifiedpatients <- as.data.frame(cbind(c("PH1", "PH1", "PH1", "PH1", "PH2", "PH2", "PH2", "PH2", "PH3", "PH3", "PH3", "PH3"), c("Yes", "Yes", "No", "No","Yes", "Yes", "No", "No","Yes", "Yes", "No", "No"), c("FALSE", "TRUE", "FALSE", "TRUE","FALSE", "TRUE", "FALSE", "TRUE","FALSE", "TRUE", "FALSE", "TRUE"), c(88.64865, 11.35135, 58.04196,	41.95804, 89.09091, 10.90909, 62.50000, 37.50000, 94.73684, 5.263158, 63.38290, 36.617100))) %>% mutate(V4 = as.numeric(V4))
missclassifiedpatients2 <- as.data.frame(cbind(c("PH1", "PH1", "PH1", "PH1", "PH2", "PH2", "PH2", "PH2", "PH3", "PH3", "PH3", "PH3"), c("Yes", "Yes", "No", "No","Yes", "Yes", "No", "No","Yes", "Yes", "No", "No"), c("FALSE", "TRUE", "FALSE", "TRUE","FALSE", "TRUE", "FALSE", "TRUE","FALSE", "TRUE", "FALSE", "TRUE"), c(39.70944, 10.44776, 60.29056	,	89.55224, 23.72881, 5.970149, 76.27119	, 94.029851, 17.43341	, 1.99005, 82.56659, 98.00995))) %>% mutate(V4 = as.numeric(V4))
missclassifiedpatients3 <- as.data.frame(cbind(c("PH1", "PH1", "PH1", "PH1", "PH2", "PH2", "PH2", "PH2", "PH3", "PH3", "PH3", "PH3"), c("Yes", "Yes", "No", "No","Yes", "Yes", "No", "No","Yes", "Yes", "No", "No"), c("FALSE", "TRUE", "FALSE", "TRUE","FALSE", "TRUE", "FALSE", "TRUE","FALSE", "TRUE", "FALSE", "TRUE"), c(26.71010	, 3.420195, 40.55375	,	29.315961, 15.96091, 1.954397, 51.30293	, 30.781759, 11.72638	, 0.6514658, 55.53746, 32.0846906))) %>% mutate(V4 = as.numeric(V4))


missclassifiedpatients %>% ggplot(aes(x = V2, y = V4, fill = V3)) + geom_col() + facet_wrap(~V1) + scale_fill_manual(values = c("deeppink","midnightblue")) + theme_minimal() + labs(x = "Diagnostically challenging patients", y = "Percentage")  + guides(fill=guide_legend(title="Treatment \nstatus"))

missclassifiedpatients2 %>% ggplot(aes(x = V3, y = V4, fill = V2)) + geom_col() + facet_wrap(~V1) + scale_fill_manual(values = c("deeppink","midnightblue")) + theme_minimal()  + labs(x = "Treatment status", y = "Percentage")  + guides(fill=guide_legend(title="Diagnostically \nchallenging \npatients"))

missclassifiedpatients3 %>% ggplot(aes(x = V3, y = V4, fill = V2)) + geom_col() + facet_wrap(~V1) + scale_fill_manual(values = c("deeppink","midnightblue")) + theme_minimal()  + labs(x = "Treatment status", y = "Percentage")  + guides(fill=guide_legend(title="Diagnostically \nchallenging \npatients"))

```

```{r}
table(ClusterphenoFPFN$cluster6, ClusterphenoFPFN$PH1FPFNbinary) %>% kable(., caption = "PH1 vs PH2 & PH3") %>% kable_styling(full_width = TRUE)
table(ClusterphenoFPFN$cluster6, ClusterphenoFPFN$PH2FPFNbinary) %>% kable(., caption = "PH2 vs PH1 & PH3") %>% kable_styling(full_width = TRUE)
table(ClusterphenoFPFN$cluster6, ClusterphenoFPFN$PH3FPFNbinary) %>% kable(., caption = "PH3 vs PH1 & PH2") %>% kable_styling(full_width = TRUE)
```

```{r, cache = FALSE}
save.image("Preprocessing.RData")
#load("Preprocessing.RData")
```







